### 2021-04-16 paper/maxdet.tex
Started paper with introduction of the xccy model.

\documentclass[12pt, oneside]{article} 
\usepackage[a4paper]{geometry}              
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[table]{xcolor}

%SetFonts
\usepackage[T1]{fontenc}
\usepackage[bitstream-charter]{mathdesign}
%SetFonts

%define example environment 
\newcounter{examplecounter}
\newenvironment{example}%
{%
\small\begin{quote}%
\refstepcounter{examplecounter}%
\textbf{Example \arabic{examplecounter}}%
\quad%
}%
{%\schluss%
\end{quote}%
}

%define remark environment 
\newcounter{remarkcounter}
\newenvironment{remark}%
{%
\small\begin{quote}%
\refstepcounter{remarkcounter}%
\textbf{Remark \arabic{remarkcounter}}%
\quad%
}%
{%\schluss%
\end{quote}%
}

%define remark environment 
\newcounter{questioncounter}
\newenvironment{question}%
{%
\small\begin{quote}%
\refstepcounter{questioncounter}%
\textbf{Question \arabic{questioncounter}}%
\quad%
}%
{%\schluss%
\end{quote}%
}

%define important environment
\newenvironment{important}{\begin{quote}%
\textbf{Important:}%
\quad
}{%
\end{quote}%
}

\newcommand{\qed}{\nobreak \ifvmode \relax \else
      \ifdim\lastskip<1.5em \hskip-\lastskip
      \hskip1.5em plus0em minus0.5em \fi \nobreak
      \vrule height0.5em width0.5em depth0.25em\fi}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}{Definition}
\newenvironment{proof}[1][Proof]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}

\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\fol}{\mathcal{F}}
\newcommand{\one}{\mathbb{1}}
\newcommand{\hest}{\textsc{Hesten}}
\newcommand{\texp}{T_{\rm exp}}

\usepackage{lettrine}

\begin{document}
\noindent{\Huge \textbf\textsf{{Completing the correlation matrix in a hybrid model.}}}

\noindent \textit{by Horst Köhler, Thomas Streuer, Olaf Dreyer}

\vspace{.5cm}

\section{Introduction}

\section{The cross currency model}
To make the issues more concrete we introduce the basic cross currency model as an example. This model consists of two interest rate models and a model for the exchange rate. We will denote the base currency by $E$ and assume that we model a rate $x_E$ together with a volatility $\nu_E$. The details of the model are not important since we are only concerned about the correlation of the stochastic processes. An example of such a model would be a Heath-JarrowÂ€-Morton model with a stochastic volatility. The model for the foreign currency $A$ will also consist of a model for the rate $x_A$ and the volatility $\nu_A$. The remaining model will be for the exchange rate 
\begin{equation}
	X_E^A,
\end{equation}
where the lower index $E$ denotes the currency we are converting into and the upper index is the currency we are converting from. If $m_A$ is an amount in currency $A$ then
\begin{equation}
	X_E^A m_A
\end{equation}
is the corresponding amount in the currency $E$. We also assume that we are modeling the volatility $\nu_{EA}$ of the exchange rate. An example of such a model would be a Heston model \cite{heston}. Altogether we thus have these six stochastic variables:
\begin{equation}
	x_E, \nu_E, x_A, \nu_A, X_E^A, \nu_{EA}
\end{equation}
Let us denote this set of variables by $I$. For two elements $a,b\in I$ let 
\begin{equation}
	(a,b) = \langle W_a, W_b\rangle
\end{equation}
be the correlation coefficient\footnote{We introduce this non-standard notation here to improve the readability in later chapters.} for the stochastic processes of $a$ and $b$. For all $a,b\in I$ we have
\begin{align}
	(a,b) & = (b,a) \\
	(a,a) & = 1\\
	\vert(a,b)\vert & \le 1
\end{align}
In principle one could imagine that all entries of $W$ are determined in one calibration using an assortment of financial products. This is usually not what one does though because the number of parameters is too large and hence the calibration is too unstable but also because one will have already performed calibrations for the interest models themselves and thus
\begin{equation}
	(x_E, \nu_E)\text{\ and\ }(x_A, \nu_A)
\end{equation}
are already known. In addition to these two, one might decide to calibrate the coefficients
\begin{equation}
	(x_E, X_E^A),(x_E, x_A), (x_A, X_E^A)\text{, and\ } (X_E^A, \nu_{EA}).
\end{equation}
The question that arises is then how to determine the remaining coefficients (see figure \ref{fig.xccymodel} of the correlation matrix.

\begin{figure}[hbt]
  \begin{center}
  	\includegraphics[width=10cm]{img/xccymodel.png}
  \end{center}
  \caption{The correlation matrix for the cross currency model. The coefficients $(x_E, \nu_E)$ and $(x_A, \nu_A)$ (the boxes with the solid lines) will be determined in two separate interest rate calibrations. The coefficients $(x_E, X_E^A)$, $(x_E, x_A)$, $(x_A, X_E^A)$, and $(X_E^A, \nu_{EA})$ (the boxes with the dashed lines) will be determined in the cross currency calibration. The question then arises what to do witht he remaining coefficients?}\label{fig.xccymodel}
\end{figure}

\section{Independence, maximal determinant, and entropy}
Before we answer the question of the last section in full generality we look at a simpler case.

\section{The completion procedure}

\section{The general hybrid model}

\begin{thebibliography}{MM}
	\bibitem{heston} Steven L. Heston, A Closed-Form Solution for Options with Stochastic Volatility with Applications to Bond and Currency Options, The Review of Financial Studies, \textbf{6}(2), April 1993, 327--343.
\end{thebibliography}

\end{document}

### 2021-04-16 paper/maxdet.tex
Discussion on why we maximize the determinant.

@@
 \end{equation}
-The question that arises is then how to determine the remaining coefficients (see figure \ref{fig.xccymodel} of the correlation matrix.
+The question that arises is then how to determine the remaining coefficients (see figure \ref{fig.xccymodel}) of the correlation matrix.
 
@@
 The coefficients $(x_E, X_E^A)$, $(x_E, x_A)$, $(x_A, X_E^A)$, and $(X_E^A, \nu_{EA})$ (the boxes with the dashed lines) will be determined in the cross currency calibration.
-The question then arises what to do witht he remaining coefficients?}\label{fig.xccymodel}
+The question then arises what to do with he remaining coefficients?}\label{fig.xccymodel}
 \end{figure}
@@
 \section{Independence, maximal determinant, and entropy}
-Before we answer the question of the last section in full generality we look at a simpler case.
+\subsection{A simpler problem}
+Before we answer the question of the last section in full generality we look at a simpler case to determine how we want to complete the matrix $M$.
+As in the previous section we denote by $I$ be the set of stochastic processes.
+For this simple case we assume that there are two sets $\alpha, \beta\subset I$ with $\alpha\cup\beta = I$ and $\alpha\cap\beta = \gamma\ne \emptyset$.
+For any $\delta\subset I$ we will denote by
+\begin{equation}
+	M[\delta]
+\end{equation}
+the principal submatrix of $M$ with row and column indices in $\delta$. In general we will denote for $\delta,\eta\subset I$ by 
+\begin{equation}
+	M[\delta,\eta]
+\end{equation}
+the submatrix of $M$ with row indices in $\delta$ and column indices in $\eta$. We assume that
+\begin{equation}
+	M[\alpha] \text{\ and\ } M[\beta] 
+\end{equation}
+are known. The matrix $M$ thus has the form
+\begin{equation}\label{eqn.corrmatrix}
+	M = \begin{pmatrix}
+		C & D & W \\
+		D^T & E & F \\
+		W^T & F^T & G
+	\end{pmatrix}.
+\end{equation}
+Using the notation that we have just introduced we have
+\begin{align}
+	M[\alpha] & = \begin{pmatrix}
+		C & D  \\
+		D^T & E 
+	\end{pmatrix} \\
+	M[\beta] & = \begin{pmatrix}
+		E & F  \\
+		F^T & G 
+	\end{pmatrix} \\
+	M[\gamma] & = E \\
+	M[ \alpha-\gamma, \alpha-\gamma] & = C \\
+	M[ \alpha-\gamma, \gamma] & = D \\
+	M[ \alpha-\gamma, \beta-\gamma] & = W \\
+	M[ \gamma, \beta-\gamma] & = F \\
+	M[ \beta-\gamma, \beta-\gamma] & = F
+\end{align}
+The part of the matrix that we need to determine is $W$.
+If the overlap $E$ didn't exist we would just set $W=0$ and thus ensure that $\alpha$ and $\beta$ are independent.
+This is the desirable result because any other choice would assume information about the correlation between $\alpha$ and $\beta$ that we do not have.
+Since the overlap $E$ is present, we have to make a different choice.
+
+\subsection{Conditioning \& Schur complement}
+For $\delta\subset I$ let us assume that $M[\delta]$ is invertible (we will weaken this assumption later). Then let 
+\begin{equation}
+	M/M[\delta] = M[I-\delta,I-\delta] - M[I-\delta, \delta]M[\delta]^{-1}M[\delta, I-\delta]
+\end{equation}
+be the Schur complement\cite{hayns} (see \cite{horn} for a an overview of the properties of the Schur complement) of $M[\delta]$ in $M$.
+In our context the Schur complement arises naturally when we ask for the correlation matrix of a subset of the stochastic variables in $I$ conditioned on the values of the remaining variables.
+If the variables that we are conditioning on are in $\gamma$ then $M/M[\gamma]$ is the correlation matrix for the variables in $I-\gamma$(for a proof see \cite[chapter 8]{rao} and \cite{cottle}).
+
+\subsection{The completion}
+With the result from the last section we have a way to proceed.
+We can look at the correlation matrix for the elements in $I$ that we obtain when we condition on the elements in $\gamma=\alpha\cap\beta$.
+This matrix is given by the Schur complement
+\begin{equation}
+	M/M[\sigma].
+\end{equation}
+We can use equation (\ref{eqn.corrmatrix}) to write
+\begin{align}
+	M/M[\sigma] & = M/E \\
+	& = \begin{pmatrix}
+		C - D E^{-1} D^T & W - DE^{-1}F \\
+		W^T - F^TE^{-1}D^T & G - F^TE^{-1}F
+		\end{pmatrix}.
+\end{align}
+A special choice for $W$ now presents itself. If we choose 
+\begin{equation}\label{eqn.choice}
+	W = DE^{-1}F 
+\end{equation}
+the off-diagonal matrices vanish and the matrix becomes block-diagonal. We can even interpret the remaining blocks. We note that 
+\begin{align}
+	M[\alpha]/M[\gamma] & = C - D E^{-1} D^T \\
+	M[\beta]/M[\gamma] & = G - F^TE^{-1}F.
+\end{align}
+Again, using the result from the previous section, we see that the remaining blocks give the correlation matrix for $\alpha$ and $\beta$ respectively, when we condition on $\gamma$.
+The block-diagonal form in equation (\ref{eqn.choice}) says that the variables $\alpha$ conditioned on $\gamma$ are independent from the variables $\beta$ conditioned on $\gamma$.
+If $f$ is the corresponding density function we have
+\begin{equation}
+	f(x_{I-\gamma} \vert \gamma ) = f(x_{\alpha-\gamma} \vert \gamma )f(x_{\beta-\gamma} \vert \gamma ).
+\end{equation}
+Given that $\gamma$ is shared by both $\alpha$ and $\beta$ this is the most independence that we can hope for.
+
+\subsection{Maximal determinant}
+Another way to characterize the choice of $W$ in equation (\ref{eqn.choice}) is by looking at the determinant of $M$. In general we have 
+\begin{align}
+	\det(M) & = \det(M[\gamma]) \det(M/M[\gamma]) \\
+	& \le \det(M[\gamma]) \det(M[\alpha]/M[\gamma]) \det(M[\beta/M[\gamma]),\label{eqn.upperbound}
+\end{align}
+(see \cite{horn} for the first equality), where we have used the Fischer inequality in the second line.
+With $W$ as in equation (\ref{eqn.choice}) the Schur complement $M/M[\gamma]$ becomes block-diagonal and we obtain
+\begin{equation}
+	\det(M) = \det(M[\gamma]) \det(M[\alpha]/M[\gamma]) \det(M[\beta/M[\gamma]).
+\end{equation}
+The determinant thus reaches the upper bound of equation (\ref{eqn.upperbound}). Choosing $W$ as in equation (\ref{eqn.choice}) we thus maximize the determinant of $M$.
+
+\subsection{Zeroes \& the inverse matrix}
+To state the next characterization of $W$ it is convenient to choose a different order for the rows and columns of $M$. Let us introduce the matrices
+\begin{align}
+	A = M[I-\gamma] \\
+	B = M[\gamma, I - \gamma]
+\end{align}
+and write
+\begin{equation}
+	M = \begin{pmatrix}
+		E & B \\
+		B^T & A
+	\end{pmatrix}.
+\end{equation}
+Babachiewicz \cite{horn} found a way to write the inverse of $M$ in terms of the Schur complement:
+\begin{equation}
+	M^{-1} = \begin{pmatrix}
+		E^{-1} + E^{-1}B(M/E)^{-1}B^TE^{-1} & - E^{-1}B (M/E)^{-1} \\
+		-(M/E)^{-1}B^TE^{-1} & (M/E)^{-1}
+	\end{pmatrix}.
+\end{equation}
+With our choice of $W$, $M/E$ is block-diagonal. It follows that $(M/E)^{-1}$ is also block-diagonal:
+\begin{equation}
+	(M/E)^{-1} = \begin{pmatrix}
+		(M[\alpha]/M[\gamma])^{-1} & 0 \\
+		0 & (M[\beta]/M[\gamma])^{-1}
+	\end{pmatrix}
+\end{equation}
+We see that the inverse of $M$ has zeroes in exactly the places where the undetermined matrix $W$ used to be.
+
+\subsection{Entropy}
+The determinant of $M$ is directly related to the entropy of the distribution that is described by $M$:
+\begin{equation}
+	S = \frac{1}{2}\big(\log(\det(M)) + n \log(2\pi) + n\big)
+\end{equation}
+Because we are maximizing the determinant we are also maximizing the entropy.
+
+\subsection{The semidefinite case}
 
@@
 Heston, A Closed-Form Solution for Options with Stochastic Volatility with Applications to Bond and Currency Options, The Review of Financial Studies, \textbf{6}(2), April 1993, 327--343.
+	\bibitem{hayns} Emilie Virginia Haynsworth, Determination of the inertia of a partitioned Hermitian matrix,  Linear  Algebra and its Applications \textbf{1}, 1968, 73--81. 
+	\bibitem{horn} Roger A. Horn, Basic  Properties  of  the  Schur  Complement. In Fuzhen Zhang (Ed.) The Schur complement and its application (17--46). Springer, 2005.
+	\bibitem{rao} C. Radharkishna Rao, Linear Statistical Inference and its Applications, 2nd edition, John Wiley \& Sons, Inc, 2002.
+	\bibitem{cottle} Richard W. Cottle, Manifestations of the Schur complement, Linear Algebra and its Applications \textbf{8}, 1974, 189--211. 
 \end{thebibliography}

### 2021-04-17 paper/maxdet.tex
Explicit form of M^{-1} added.

@@
 \begin{equation}
-	M/M[\sigma].
+	M/M[\gamma].
 \end{equation}
@@
 \begin{align}
-	M/M[\sigma] & = M/E \\
+	M/M[\gamma] & = M/E \\
 	& = \begin{pmatrix}
@@
 \end{equation}
-Given that $\gamma$ is shared by both $\alpha$ and $\beta$ this is the most independence that we can hope for.
+Given that $\gamma$ is shared by both $\alpha$ and $\beta$ this is the most independence we can hope for.
 
@@
 \subsection{Zeroes \& the inverse matrix}
-To state the next characterization of $W$ it is convenient to choose a different order for the rows and columns of $M$. Let us introduce the matrices
-\begin{align}
-	A = M[I-\gamma] \\
-	B = M[\gamma, I - \gamma]
-\end{align}
-and write
+A remarkable formula due to Babachiewicz \cite{horn} expresses the inverse of a matrix in terms of the Schur complement. For a matrix
 \begin{equation}
-	M = \begin{pmatrix}
-		E & B \\
-		B^T & A
+	\Sigma = \begin{pmatrix}
+		\Sigma_{11} & \Sigma_{12} \\ 
+		\Sigma_{21} & \Sigma_{22}
+	\end{pmatrix}
+\end{equation}
+we have
+\begin{equation}
+	\Sigma^{-1} = \begin{pmatrix}
+		\Sigma_{11}^{-1} + \Sigma_{11}^{-1}\Sigma_{12}(\Sigma/\Sigma_{11})^{-1}\Sigma_{21}\Sigma_{11}^{-1} &-\Sigma_{11}^{-1} \Sigma_{12} (\Sigma/\Sigma_{11})^{-1} \\ 
+		-(\Sigma/\Sigma_{11})^{-1}\Sigma_{21}\Sigma_{11}^{-1} & (\Sigma/\Sigma_{11})^{-1}
 	\end{pmatrix}.
 \end{equation}
-Babachiewicz \cite{horn} found a way to write the inverse of $M$ in terms of the Schur complement:
+This formula has the remarkable property that only the inverses 
+\begin{equation}
+	\Sigma_{11}^{-1} \text{\ and\ } (\Sigma/\Sigma_{11})^{-1}
+\end{equation}
+appear in it. With our choice of $W$ in equation (\ref{eqn.choice}) the Schur complement $M/E$ is block diagonal. This means that also its inverse is block diagonal. Indeed, we have
+\begin{equation}
+	(M/E)^{-1} = \begin{pmatrix}
+		(M[\alpha]/E)^{-1} & 0 \\
+		0 & (M[\beta]/E)^{-1}
+	\end{pmatrix}.
+\end{equation}
+This allows us to write the inverse of $M$ explicitly:
 \begin{equation}
 	M^{-1} = \begin{pmatrix}
-		E^{-1} + E^{-1}B(M/E)^{-1}B^TE^{-1} & - E^{-1}B (M/E)^{-1} \\
-		-(M/E)^{-1}B^TE^{-1} & (M/E)^{-1}
-	\end{pmatrix}.
+		(M[\alpha]/E)^{-1} & -(M[\alpha]/E)^{-1} D E^{-1} & 0 \\
+		- E^{-1} D^T (M[\alpha]/E)^{-1} & \Xi & - E^{-1} F (M[\beta]/E)^{-1} \\
+		0 & - (M[\beta]/E)^{-1} F^T E^{-1} & (M[\beta]/E)^{-1}
+	\end{pmatrix},
 \end{equation}
-With our choice of $W$, $M/E$ is block-diagonal. It follows that $(M/E)^{-1}$ is also block-diagonal:
+with
 \begin{equation}
-	(M/E)^{-1} = \begin{pmatrix}
-		(M[\alpha]/M[\gamma])^{-1} & 0 \\
-		0 & (M[\beta]/M[\gamma])^{-1}
-	\end{pmatrix}
+	\Xi = E^{-1} + E^{-1}\left(D^T(M[\alpha]/E)^{-1}D + F (M[\beta]/E)^{-1} F^T\right)E^{-1}.
 \end{equation}
-We see that the inverse of $M$ has zeroes in exactly the places where the undetermined matrix $W$ used to be.
+We see that the inverse of $M$ has zeroes in exactly those places where the undetermined matrix $W$ used to be.
 

### 2021-04-17 paper/maxdet.tex
Small typos corrected

@@
 \end{equation}
-the principal submatrix of $M$ with row and column indices in $\delta$. In general we will denote for $\delta,\eta\subset I$ by 
+the principal submatrix of $M$ with row and column indices in $\delta$. For two sets $\delta,\eta\subset I$ we denote by 
 \begin{equation}
@@
 \end{equation}
-are known. The matrix $M$ thus has the form
+are known. The matrix $M$ can then be written in the form
 \begin{equation}\label{eqn.corrmatrix}
@@
 		W^T & F^T & G
-	\end{pmatrix}.
+	\end{pmatrix},
 \end{equation}
-Using the notation that we have just introduced we have
+where we make the following identifications:
 \begin{align}

### 2021-04-19 paper/maxdet.tex
Added the first theorem for the semidefinite case

@@
 	M[ \gamma, \beta-\gamma] & = F \\
-	M[ \beta-\gamma, \beta-\gamma] & = F
+	M[ \beta-\gamma, \beta-\gamma] & = G
 \end{align}
@@
 
-\subsection{Conditioning \& Schur complement}
+\subsection{Conditioning \& the Schur complement}
 For $\delta\subset I$ let us assume that $M[\delta]$ is invertible (we will weaken this assumption later). Then let 
@@
 \end{equation}
-The determinant thus reaches the upper bound of equation (\ref{eqn.upperbound}). Choosing $W$ as in equation (\ref{eqn.choice}) we thus maximize the determinant of $M$.
+The determinant thus reaches the upper bound of equation (\ref{eqn.upperbound}). Choosing $W$ as in equation (\ref{eqn.choice}) thus maximizes the determinant of $M$.
 
@@
 A remarkable formula due to Babachiewicz \cite{horn} expresses the inverse of a matrix in terms of the Schur complement. For a matrix
-\begin{equation}
+\begin{equation}\label{eqn.sigma}
 	\Sigma = \begin{pmatrix}
@@
 \subsection{The semidefinite case}
+In the previous sections we have assumed that $E=M[\gamma]$ is non-singular so that $E^{-1}$ exists.
+In this section we will only assume that $E$ is positive semi-definite.
+It is in this case that our results differ from those in \cite{smith}.
+Because of this we give new proves for the following results that do not make a distinction between the definite and semi-definite case.
+We start by noting some properties of symmetric matrices that are also positive semi-definite.
+
+\begin{lemma}\label{lem.symm}
+	Let $\Sigma$ be a symmetric matrix that is partitioned as in equation (\ref{eqn.sigma}). If $\Sigma$ is positive semidefinite then we have
+	\begin{enumerate}
+		\item $\mathcal{C}(\Sigma_{12})\subset \mathcal{C}(\Sigma_{11})$ \\
+		\item $\mathcal{R}(\Sigma_{21})\subset \mathcal{R}(\Sigma_{11})$ \\
+		\item $r(\Sigma) = r(\Sigma_{11}) + r(\Sigma_{22})$
+	\end{enumerate}
+	Here $\mathcal{C}(\cdot)$ denotes the span of the column vectors of a matrix, $\mathcal{R}(\cdot)$ denotes the row space, and $r(\cdot)$ denotes the rank of a matrix.  
+\end{lemma}
+
+\begin{proof}
+	The proof is given in appendix \ref{app.symm}. \qed
+\end{proof}
+
+In the following we no longer assume that the matrices are non-singular.
+It turns out that we can still define the Schur complement of a matrix.
+For a matrix $E$ we will denote by $E^-$ a generalized inverse of $E$.
+$E^-$ is a matrix that satisfies
+\begin{equation}
+	E E^{-}E = E.
+\end{equation}
+A particular example of a generalized inverse is the Moore-Penrose inverse $E^+$ \cite{roger}.
+For now, we will not assume that we are dealing with $E^+$ but keep the discussion general.
+Given a particular choice of generalized inverse we define
+\begin{equation}
+	\Sigma/\Sigma_{11} = \Sigma_{22} - \Sigma_{21}\Sigma_{11}^-\Sigma_{12}.
+\end{equation}
+If $\Sigma$ is a symmetric matrix that is also positive semi-definite it turns out that the Schur complement does not depend on the choice of the particular generalized inverse used in its definition.
+
+\begin{lemma} Let $\Sigma$ be a symmetric matrix that is partitioned as in equation (\ref{eqn.sigma}). If $\Sigma$ is positive semidefinite then
+\begin{equation}
+	\Sigma/\Sigma_{11} = \Sigma_{22} - \Sigma_{21}\Sigma_{11}^-\Sigma_{12}
+\end{equation}	
+does not depend on the choice of the generalized inverse $\Sigma_{11}^-$.
+\end{lemma}
+
+\begin{proof}
+	The proof is given in appendix \ref{app.symm}. \qed
+\end{proof}
+
+The next theorem characterizes positive symmetric matrices in terms of their Schur complement:
+
+\begin{theorem}\label{theo.charact}
+	Let $\Sigma$ be a symmetric matrix that is partitioned as in equation (\ref{eqn.sigma}) where $\Sigma_{11}$ and $\Sigma_{22}$ are square.
+Then $\Sigma\ge 0$ if and only if $\Sigma_{11}\ge 0$, $\mathcal{C}(\Sigma_{12})\subset \mathcal{C}(\Sigma_{11})$, and $\Sigma/\Sigma_{11}\ge 0$.
+\end{theorem}
+
+\begin{proof}
+	See \cite[Theorem 1.20]{horn} and \cite{albert}. \qed
+\end{proof}
+
+In equation (\ref{eqn.choice}) we defined $W$ using the inverse $E^{-1}$.
+We are now going to define $W$ using a general inverse instead.
+The next theorem shows that this again defines a positive semidefinite completion of $M$.
+The following theorems then show that the other properties also carry over to the semidefinite case.
+
+\begin{theorem}
+	Let $M$ be as in equation (\ref{eqn.corrmatrix}). Let both $M[\alpha]$ and $M[\beta]$ be positive semidefinite. Setting 
+	\begin{align}
+		W & = DE^-F \\
+		  & = M[ \alpha-\gamma, \gamma]M[\gamma]^-M[ \gamma, \beta-\gamma]		
+	\end{align}
+	turns $M$ into a positive semidefinite matrix.
+\end{theorem}
+
+\begin{proof}
+	Since both $M[\alpha]$ and $M[\beta]$ are positive semidefinite it follows from theorem \ref{theo.charact} that
+	\begin{align}
+		E & = M[\gamma] \ge 0 \\
+		\mathcal{C}(D^T) & = \mathcal{C}(M[ \alpha-\gamma, \gamma]^T) \subset \mathcal{C}(E) \\
+		\mathcal{C}(F) & = \mathcal{C}(M[ \gamma, \beta-\gamma ]) \subset \mathcal{C}(E) \\
+		M[\alpha] / E & \ge 0 \\
+		M[\beta] / E & \ge 0
+	\end{align}
+	From this it follows that
+	\begin{align}
+		E & = M[\gamma] \ge 0 \\
+		\mathcal{C}(D^T\ F) & = \mathcal{C}( M[\gamma, I - \gamma] ) \subset  \mathcal{C}(E) \\
+		M/E & = \begin{pmatrix}
+			M[\alpha] / E & 0 \\
+			0 & M[\beta] / E
+		\end{pmatrix} \ge 0
+	\end{align}
+	Using theorem \ref{theo.charact} in the other direction now yields the result. \qed
+\end{proof}
 
@@
 	\bibitem{cottle} Richard W. Cottle, Manifestations of the Schur complement, Linear Algebra and its Applications \textbf{8}, 1974, 189--211. 
+	\bibitem{smith} Smith, R. (2008). The positive definite completion problem revisited Linear Algebra and its Applications  429(7), 1442-1452. https://dx.doi.org/10.1016/j.laa.2008.04.020
+	\bibitem{ouellette} Ouellette, D. (1981). Schur complements and statistics Linear Algebra and its Applications  36, 187-295. https://dx.doi.org/10.1016/0024-3795(81)90232-9
+	\bibitem{roger} Penrose, R. (1955). A generalized inverse for matrices Mathematical Proceedings of the Cambridge Philosophical Society  51(3), 406-413. https://dx.doi.org/10.1017/s0305004100030401
+	\bibitem{albert} Albert, A.
+(1969).
+Conditions for Positive and Nonnegative Definiteness in Terms of Pseudoinverses SIAM Journal on Applied Mathematics  17(2), 434-440.
+https://dx.doi.org/10.1137/0117041
 \end{thebibliography}
 
+\appendix
+
+\section{Symmetric matrices \& the Schur complement}
+\label{app.symm}
+
 \end{document}

### 2021-04-20 paper/maxdet.tex
Added the uniqueness result.

@@
 \newtheorem{theorem}{Theorem}[section]
+\newtheorem{corollary}{Corollary}[section]
 \newtheorem{lemma}[theorem]{Lemma}
@@
 we have
-\begin{equation}
+\begin{equation}\label{eqn.baba}
 	\Sigma^{-1} = \begin{pmatrix}
@@
 For now, we will not assume that we are dealing with $E^+$ but keep the discussion general.
+Let us look at a general matrix $\Sigma$ as in equation (\ref{eqn.sigma}).
 Given a particular choice of generalized inverse we define
@@
 \begin{theorem}
-	Let $M$ be as in equation (\ref{eqn.corrmatrix}). Let both $M[\alpha]$ and $M[\beta]$ be positive semidefinite. Setting 
+	Let $M$ be as in equation (\ref{eqn.corrmatrix}). Let both $M[\alpha]$ and $M[\beta]$ be positive semidefinite. Let $E^-$ be a generalized inverse of $E$ that is also symmetric. Setting 
 	\begin{align}
@@
 
+We next look at the generalized inverse $M^+$ of $M$.
+It turns out that it can be uniquely expressed in terms of the Schur complement.
+This is a generalization of the Babachiewicz's result from equation (\ref{eqn.baba}) to the semidefinite case.
+
+\begin{theorem}
+	Let $\Sigma$ be a symmetric matrix partitioned as in equation (\ref{eqn.sigma}). Then we have
+	\begin{equation}
+		\Sigma^+ = \begin{pmatrix}\label{eqn.mpinverse}
+		\Sigma_{11}^{+} + \Sigma_{11}^{+}\Sigma_{12}(\Sigma/\Sigma_{11})^{+}\Sigma_{21}\Sigma_{11}^{+} &-\Sigma_{11}^{+} \Sigma_{12} (\Sigma/\Sigma_{11})^{+} \\ 
+		-(\Sigma/\Sigma_{11})^{+}\Sigma_{21}\Sigma_{11}^{+} & (\Sigma/\Sigma_{11})^{+}
+	\end{pmatrix}
+	\end{equation}
+	if and only if 
+	\begin{equation}\label{eqn.rank}
+			r(\Sigma) = r(\Sigma_{11}) + r(\Sigma_{22}).		
+	\end{equation}
+\end{theorem}
+
+\begin{proof}
+	See \cite[section 4]{ouellette} and \cite{horn}. \qed
+\end{proof}
+
+For a symmetric positive semidefinite matrix $\Sigma$ the condition on the rank of $\Sigma$ is automatically satisfied. We thus have the following corollary:
+
+\begin{corollary}\label{cor.inverse}
+	Let $\Sigma$ be symmetric and positive semidefinite. Then $\Sigma^+$ is given by equation (\ref{eqn.mpinverse}).
+\end{corollary}
+
+We can now state a unique property of our completion:
+
+\begin{theorem}
+	The completion defined by setting $W=DE^+F$ is the unique completion that turns $M$ into a symmetric positive semidefinite matrix and for which $M^+$ has zeroes in the places where $W$ used to be.
+\end{theorem}
+
+The uniqueness result stated in \cite{smith} gives this result only for the positive definite case. This result shows that it is indeed also true for the positive semidefinite case.
+
+\begin{proof}
+	Because of the previous corollary \ref{cor.inverse} we see that for any completion $W$ that turns $M$ into a symmetric positive semidefinite matrix the Moore-Penrose inverse is given by equation (\ref{eqn.mpinverse}).
+This inverse has zeroes in exactly the places where $W$ used to be if and only if $W=DE^+F$.
+\qed
+\end{proof}
+
 \section{The completion procedure}

### 2021-04-26 notes/fischer.tex
Created a note: Fischer's inequality.

\documentclass[12pt, oneside]{article} 
\usepackage[a4paper]{geometry}              
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[table]{xcolor}

%SetFonts
\usepackage[T1]{fontenc}
\usepackage[bitstream-charter]{mathdesign}
%SetFonts

%define example environment 
\newcounter{examplecounter}
\newenvironment{example}%
{%
\small\begin{quote}%
\refstepcounter{examplecounter}%
\textbf{Example \arabic{examplecounter}}%
\quad%
}%
{%\schluss%
\end{quote}%
}

%define remark environment 
\newcounter{remarkcounter}
\newenvironment{remark}%
{%
\small\begin{quote}%
\refstepcounter{remarkcounter}%
\textbf{Remark \arabic{remarkcounter}}%
\quad%
}%
{%\schluss%
\end{quote}%
}

%define remark environment 
\newcounter{questioncounter}
\newenvironment{question}%
{%
\small\begin{quote}%
\refstepcounter{questioncounter}%
\textbf{Question \arabic{questioncounter}}%
\quad%
}%
{%\schluss%
\end{quote}%
}

%define important environment
\newenvironment{important}{\begin{quote}%
\textbf{Important:}%
\quad
}{%
\end{quote}%
}

\newcommand{\qed}{\nobreak \ifvmode \relax \else
      \ifdim\lastskip<1.5em \hskip-\lastskip
      \hskip1.5em plus0em minus0.5em \fi \nobreak
      \vrule height0.5em width0.5em depth0.25em\fi}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}{Definition}
\newenvironment{proof}[1][Proof]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}

\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\range}{\text{\normalfont range}}
\newcommand{\rank}{\text{\normalfont rank}}
\newcommand{\fol}{\mathcal{F}}
\newcommand{\one}{\mathbb{1}}
\newcommand{\hest}{\textsc{Hesten}}
\newcommand{\texp}{T_{\rm exp}}
\newcommand{\detp}{{\det}_+}

\usepackage{lettrine}

\begin{document}
\noindent{\Huge \textbf\textsf{{The generalized determinant and the Fischer's inequality}}}

\noindent \textit{by Horst Köhler, Thomas Streuer, Olaf Dreyer}

\vspace{.5cm}

\section{Definition}
We begin with the definition of the generalized determinant:

\begin{definition}
	Let $A\in M_n(\F)$ ($\F = \R$, or $\C$). Let $\lambda_1, \ldots, \lambda_n$ be the eigenvalues of $A$. Let $\lambda_i=0$ for $n\ge i>r\ge 0$. Then we call the product of the nonzero eigenvalues of $A$,
	\begin{equation}
		\detp A = \prod_{i=1}^r \lambda_i,
	\end{equation}  
	the generalized determinant of $A$. In the case of $r=0$ we set $\detp A = 1$.
\end{definition}

\begin{remark}
	If $A$ is nonsingular we have $\detp A = \det A$.
\end{remark}

\begin{remark}
	The generalized determinant does not share many of the properties with the usual determinant. While the determinant is a continuous function of $A$, the generalized determinant is not. This can be seen in the following simple example:
	\begin{align}
		1 & = \detp \begin{pmatrix}
			0 & 0 \\
			0 & 1
		\end{pmatrix} \\
		& \ne \lim_{\epsilon\rightarrow 0} \detp \begin{pmatrix}
			\epsilon & 0 \\
			0 & 1
		\end{pmatrix} \\
		& = 0
	\end{align}
	The determinant of a product is equal to the product of the determinants:
	\begin{equation}
		\det(AB) = \det(A)\det(B)
	\end{equation} 
	This is not true for the generalized determinant:
	\begin{align}
		\detp \left(\begin{pmatrix}
			a & 0 \\
			0 & 0
		\end{pmatrix} \begin{pmatrix}
			0 & 0 \\
			0 & b
		\end{pmatrix} \right) & = \detp \begin{pmatrix}
			0 & 0 \\
			0 & 0
		\end{pmatrix} \\
		& = 1 \\
		& \ne \detp \begin{pmatrix}
			a & 0 \\
			0 & 0
		\end{pmatrix} \detp \begin{pmatrix}
			0 & 0 \\
			0 & b
		\end{pmatrix} \\
		& = a b
	\end{align}
\end{remark}

\section{Basic properties}
The generalized determinant does have some interesting properties though. For the following we will restrict ourselves to the case of a Hermitian matrix $A$. The following lemma will allow us to infer some properties of the generalized determinant through a limit argument:

\begin{lemma}
	Let $A\in M_n(\C)$ be Hermitian with rank $r$. Then
	\begin{equation}
		\detp A = \lim_{\epsilon\rightarrow 0} \frac{\det(A + \epsilon I_n)}{\epsilon^{n-r}}
	\end{equation}
\end{lemma}

\begin{proof}
	Let $\lambda_1, ..., \lambda_n$ be the eigenvalues of $A$.  Since the rank of $A$ is $r$ there are $n-r$ that are equal to zero. Let us assume that $\lambda_i>0$, for $i=1, \ldots, r$. Then we have 
	\begin{align}
		\det(A + \epsilon I_n) & = \prod_i (\lambda_i + \epsilon) \\
		 & = \epsilon^{n-r} \prod_{i=1}^r (\lambda_i + \epsilon)
	\end{align} 
	Since the generalized determinant is the product of the nonzero eigenvalues we obtain the equality by canceling the $\epsilon^{n-r}$ and taking the limit. \qed
\end{proof}

\begin{lemma}
	We have the following properties (again, $r$ is the rank of $A$):
	\begin{enumerate}
		\item For $\rho\ne 0$ we have $\detp (\rho A) = \rho^{n-r} \detp (A)$. 
		\item For $A\in M_n(\C)$ and $B\in M_m(\C)$ we have
		\begin{equation}
			\detp \begin{pmatrix}
				A & 0 \\
				0 & B 
			\end{pmatrix} = \detp A\; \detp B.
		\end{equation}
		\item For any unitary $U$ we have 
		\begin{equation}
			\detp U A U^\star = \detp A.
		\end{equation} 
	\end{enumerate}
\end{lemma}

\section{Fischer's inequality}
For a general Hermitian matrix 
\begin{equation}
	H = \begin{pmatrix}
		A & B \\
		B^\star & C 
	\end{pmatrix} \in M_{n+m}(\C),
\end{equation} 
with $A\in M_n(\C), C\in M_m(\C)$, we have
\begin{equation}
	\det H \le \det A \det C,	
\end{equation}
with equality if and only if $B=0$.

\begin{remark}\label{rem.nofischer}
	This inequality is not true for the generalized determinant. Let 
	\begin{equation}
		H = \begin{pmatrix}
			0 & 1 \\
			1 & b
		\end{pmatrix}.
	\end{equation}
	Then we have $\detp H = \det H = -1$, but 
	\begin{equation}
		\detp 0\; \detp b = b.
	\end{equation}
	Choosing $b<-1$ gives a violation of the inequality for generalized determinants.
\end{remark}

A similar inequality is valid if we further restrict the class of matrices that we are dealing with. We will show that Fischer's inequality is valid for positive semidefinite matrices. The key ingredient that makes this possible is that these matrices have the \emph{row} and \emph{column inclusion property}.

\begin{definition}
	For every $k=1, \ldots, n-1$ let $A\in M_n(\C)$ be partitioned as follows:
	\begin{equation}
	A = \begin{pmatrix}
		A_{11} & A_{12} \\
		A_{21} & A_{22}
	\end{pmatrix},
\end{equation}
with $A_{11}\in M_k(\C)$. We say that $A$ satisfies the column inclusion property if for all $k$ we have $\range(A_{12})\subset\range(A_{11})$. The matrix $A$ has the row inclusion property if $A^\star$ has the column inclusion property.
\end{definition}

\begin{remark}\label{eqn.columnincl}
	The following statements are equivalent:
	\begin{enumerate}
		\item $A$ has the column inclusion property.
		\item If $\mathcal{N}(A)$ denotes the null space of the matrix $A$, then
		\begin{equation}\label{eqn.nullincl}
			\mathcal{N}(A_{11}^\star)\subset \mathcal{N}(A_{12}^\star).
		\end{equation}
		\item Every column of $A_{12}$ is a linear combination of the columns of $A_{11}$.
		\item There is an $X\in M_{k,n-k}$ with $A_{12} = A_{11} X$.
		\item $\rank(A_{11} A_{12}) = \rank(A_{11})$.
	\end{enumerate}
\end{remark}

\begin{remark}
	The matrix in the previous remark \ref{rem.nofischer} is not positive semidefinite and it does not have the row inclusion property.
\end{remark}

\begin{lemma}
	Every positive semidefinite matrix $A$ has the row and column inclusion property.
\end{lemma}

\begin{proof}
	Let $A$ be	
	\begin{equation}
	A = \begin{pmatrix}
		A_{11} & A_{12} \\
		A_{12}^\star & A_{22}
	\end{pmatrix},
	\end{equation}
	(note that because $A$ is positive semidefinite it is also Hermitian). We will show that 
	\begin{equation}
		\mathcal{N}(A_{11})\subset \mathcal{N}(A_{12}^\star).
	\end{equation}
	(see remark \ref{eqn.columnincl}). Let $v\in \mathcal{N}(A_{11})$. Then for $w = (v,0)^T\in \C^n$ we have
	\begin{equation}
		w^\star A w = v^\star A_{11}v = 0.
	\end{equation}
	Since $A$ is positive semidefinite this implies that
	\begin{equation}
		A v = (A_{11}v, A_{12}^\star v)^T = 0,
	\end{equation}
	or $A_{12}^\star v=0$. \qed
\end{proof}

Let us note the following consequence:

\begin{lemma}
	Let $A$ be as above. Then we have
	\begin{equation}
		\rank\begin{pmatrix}
			A_{11} \\
			A_{12}^\star
		\end{pmatrix} = \rank(A_{11}).
	\end{equation}
\end{lemma}

\begin{proof}
	Because $A$ is positive semidefinite we have
	\begin{equation}\label{eqn.included}
		\mathcal{N}(A_{11})\subset \mathcal{N}(A_{12}^\star).
	\end{equation}
	We then obtain:
	\begin{align}
				\rank\begin{pmatrix}
			A_{11} \\
			A_{12}^\star
		\end{pmatrix} & = k - \dim\mathcal{N} \begin{pmatrix}
			A_{11} \\
			A_{12}^\star
		\end{pmatrix} \\
		& = k - \dim(\mathcal{N}(A_{11})\cap\mathcal{N}(A_{12}^\star))\label{eqn.cap}\\
		& = k - \dim(\mathcal{N}(A_{11})) \label{eqn.cap1}\\
		& = \rank(A_{11}).
	\end{align}
	To get to equation (\ref{eqn.cap1}) we have used equation (\ref{eqn.included}). \qed 
\end{proof}

The same result holds for $A_{22}$ and we obtain:

\begin{lemma}
	Let $A$ be as above then
	\begin{equation}
		\rank(A) \le \rank(A_{11}) + \rank(A_{12}). 
	\end{equation}
\end{lemma}

\begin{proof}
	We have
	\begin{align}
		\rank(A) & \le \rank\begin{pmatrix}
			A_{11} \\
			A_{12}^\star
		\end{pmatrix} + \rank\begin{pmatrix}
			A_{12} \\
			A_{22}
		\end{pmatrix} \\
		& = \rank(A_{11}) + \rank(A_{22}). \qed
	\end{align}
\end{proof}

From this we get the first version of the Fischer inequality:

\begin{theorem}
	Let $H\in M_n(\C)$ be a positive semidefinite matrix that is partitioned as follows:
	\begin{equation}
		H = \begin{pmatrix}
			A & B \\
			B^\star & C
		\end{pmatrix},
	\end{equation}
	with $A\in M_k(\C)$ and $1\le k< n$. Assume that 
	\begin{equation}\label{eqn.rankadd}
		\rank(H) = \rank(A) + \rank(C), 
	\end{equation}
	then
	\begin{equation}
		\detp H \le \detp A\; \detp C
	\end{equation}
\end{theorem}

\begin{proof}
	Because of equation (\ref{eqn.rankadd}) we have
	\begin{align}
		\detp H & = \lim_{\epsilon\rightarrow 0} \frac{\det(H + \epsilon I_n)}{\epsilon^{n-\rank(H)}} \\
		& \le \lim_{\epsilon\rightarrow 0} \frac{\det(A_{11} + \epsilon I_k)\det(A_{22} + \epsilon I_{n-k})}{\epsilon^{n-\rank(H)}} \\
		& = \lim_{\epsilon\rightarrow 0} \frac{\det(A_{11} + \epsilon I_k)}{\epsilon^{k-\rank(A_{11})}}\cdot\frac{\det(A_{22} + \epsilon I_{n-k})}{\epsilon^{n-k-\rank(A_{22})}} \\
		& = \detp A_{11}\; \detp A_{22}.
	\end{align}
	Which is the result. \qed
\end{proof}

\end{document}

### 2021-04-27 notes/fischer.tex
Finsihed the proof Fischer's inequality.

@@
 \begin{document}
-\noindent{\Huge \textbf\textsf{{The generalized determinant and the Fischer's inequality}}}
+\noindent{\Huge \textbf\textsf{{The generalized determinant and  Fischer's inequality}}}
 
@@
 		\end{equation}
+		\item If $\mathcal{N}(A)=\{0\}$, i.e. if $\rank(A)=n$ we have
+		\begin{equation}
+			\detp A = \det A.
+		\end{equation} 
 	\end{enumerate}
@@
 
-\begin{lemma}
+\begin{lemma}\label{lem.columnincl}
 	Every positive semidefinite matrix $A$ has the row and column inclusion property.
@@
 
+\begin{remark}
+	A simple example illustrates the importance of the condition on the ranks. Let
+	\begin{equation}
+		H = \begin{pmatrix}
+			1 & a \\
+			a & a^2
+		\end{pmatrix}.
+	\end{equation}
+	Then $\det H = 0$ and $\detp H = 1 + a^2$ (the two eigenvalues of $H$ are $0$ and $1+a^2$ with eigenvectors $(a, -1)^T$ and $(1, a)^T$ respectively). For $a>0$ we have 
+	\begin{equation}
+		\detp 1\detp a^2 = a^2.
+	\end{equation}
+	$H$ is Hermitian and positive semidefinite but Fischer's inequality does not hold for the generalized determinant. Note that 
+	\begin{equation}
+		\rank(H) = 1,
+	\end{equation}
+	while $\rank(1) + \rank(a^2) = 2$. 
+\end{remark}
+
+\section{Improvement}
+We now want ti improve the theorem by dealing with case of equality:
+
+\begin{theorem}
+	Let $H$ be Hermitian and positive semidefinite and let $H$ be partitioned as follows:
+		\begin{equation}
+		H = \begin{pmatrix}
+			A & B \\
+			B^\star & C
+		\end{pmatrix},
+	\end{equation}
+	with $A\in M_k(\C)$ and $1\le k< n$. Assume that 
+	\begin{equation}\label{eqn.rankadd}
+		\rank(H) = \rank(A) + \rank(C), 
+	\end{equation}
+	then
+	\begin{equation}
+		\detp H \le \detp A\; \detp C,
+	\end{equation}
+	with equality if and only if $B=0$.
+\end{theorem}
+
+\begin{proof}
+	Since $A$ and $C$ are Hermitian there are unitary matrices $U=U^\star\in M_k(\C)$ and $V=V^\star\in M_{n-k}(\C)$ such that
+	\begin{align}
+		U A U^\star & = \begin{pmatrix}
+			\lambda_1 & 0 & 0 \\
+			0 & \ddots & 0 \\
+			0 & 0 & \lambda_k
+		\end{pmatrix} \\
+		& \nonumber \\
+		V C V^\star & = \begin{pmatrix}
+			\mu_1 & 0 & 0 \\
+			0 & \ddots & 0 \\
+			0 & 0 & \mu_{n-k}
+		\end{pmatrix}.
+	\end{align}
+	With
+	\begin{equation}
+		W = \begin{pmatrix}
+			U & 0 \\
+			0 & V
+		\end{pmatrix}
+	\end{equation}
+	we have
+	\begin{equation}
+		W H W^\star = \begin{pmatrix}
+ 	\begin{matrix}
+			\lambda_1 &  &  \\
+			 & \ddots &  \\
+			 &  & \lambda_k
+		\end{matrix} & UBV^\star \\
+		VB^\star U^\star & \begin{matrix}
+			\mu_1 &  &  \\
+			 & \ddots &  \\
+			 &  & \mu_{n-k}
+		\end{matrix}
+ 	\end{pmatrix}
+	\end{equation}
+	We now make use of the fact that $W H W^\star$ is positive semidefinite (since $H$ is positive semidefinite).
+Let $r_A$ be the rank of $A$ and $r_C$ be the rank of $C$.
+We assume that we have ordered the eigenvalues in such a way that
+	\begin{equation}
+		\lambda_i = 0, \text{\ for\ } i>r_A,
+	\end{equation} 
+	and
+		\begin{equation}
+		\mu_j = 0, \text{\ for\ } j>r_C.
+	\end{equation}
+	Because $W H W^\star$ is positive semidefinite it follows that
+	\begin{equation}\label{eqn.lotsofzeroes}
+		(UBV^\star)_{i,j}=0, \text{\ for\ } i>r_A \text{\ or\ }j>r_C.
+	\end{equation}
+	Note the condition here. It is sufficient that either $i>r_A$ \emph{or} $j>r_C$ for $(UBV^\star)_{i,j}$ to vanish (this follows from lemma \ref{lem.columnincl}).
+
+	We now concentrate on the elements of $W H W^\star$ that are not zero. Let us define the following set of indices 
+	\begin{equation}
+		\iota = \{1,\ldots,r_A,k+1,\ldots,k+r_C\},
+	\end{equation}
+	and let us set $G = W H W^\star$. We have just established that
+	\begin{equation}
+		G[\iota^c] = 0.
+	\end{equation}
+	The remaining nonzero components are contained in $G[\iota]$. Because we are assuming that 
+	\begin{equation}
+		\rank(H) = \rank(A) + \rank(C),		
+	\end{equation}
+	and since 
+	\begin{equation}
+		\rank(G[\iota]) = \rank(H),	
+	\end{equation}
+	 $G[\iota]$ is a Hermitian positive definite matrix. Furthermore we have
+	 \begin{align}
+	 	\detp H & = \det G[\iota] \\
+	 	& \le \lambda_1\cdots\lambda_{r_A}\cdot\mu_1\cdots\mu_{r_C} \\
+	 	& = \detp A\; \detp C,
+	 \end{align}
+	 with equality if and only if the rest of $UBV^\star$ is zero. \qed
+\end{proof}
+
+\begin{remark}
+	Equation (\ref{eqn.lotsofzeroes}) is equivalent to $\mathcal{N}(C)\subset\mathcal{N}(B)$ and $\mathcal{N}(A)\subset\mathcal{N}(B^\star)$.
+These inclusions were established in lemma \ref{lem.columnincl}.
+\end{remark}
+
 \end{document}

### 2021-05-12 paper/paper_math.tex
Added the math paper. Fischer's inequality.

\documentclass[12pt, oneside]{article} 
\usepackage[a4paper]{geometry}              
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[table]{xcolor}

%SetFonts
\usepackage[T1]{fontenc}
\usepackage[bitstream-charter]{mathdesign}
%SetFonts

%define example environment 
\newcounter{examplecounter}
\newenvironment{example}%
{%
\small\begin{quote}%
\refstepcounter{examplecounter}%
\textbf{Example \arabic{examplecounter}}%
\quad%
}%
{%\schluss%
\end{quote}%
}

%define remark environment 
\newcounter{remarkcounter}
\newenvironment{remark}%
{%
\small\begin{quote}%
\refstepcounter{remarkcounter}%
\textbf{Remark \arabic{remarkcounter}}%
\quad%
}%
{%\schluss%
\end{quote}%
}

%define remark environment 
\newcounter{questioncounter}
\newenvironment{question}%
{%
\small\begin{quote}%
\refstepcounter{questioncounter}%
\textbf{Question \arabic{questioncounter}}%
\quad%
}%
{%\schluss%
\end{quote}%
}

%define important environment
\newenvironment{important}{\begin{quote}%
\textbf{Important:}%
\quad
}{%
\end{quote}%
}

\newcommand{\qed}{\nobreak \ifvmode \relax \else
      \ifdim\lastskip<1.5em \hskip-\lastskip
      \hskip1.5em plus0em minus0.5em \fi \nobreak
      \vrule height0.5em width0.5em depth0.25em\fi}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}{Definition}
\newenvironment{proof}[1][Proof]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}

\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\fol}{\mathcal{F}}
\newcommand{\one}{\mathbb{1}}
\newcommand{\rank}{\text{rank}}
\newcommand{\detp}{{\det}_+}

\usepackage{lettrine}

\begin{document}
\noindent{\Huge \textbf\textsf{{Matrix completion and semi-definite matrices.}}}

\noindent \textit{by Horst Köhler, Thomas Streuer, Olaf Dreyer}

\vspace{.5cm}

\section{Introduction}

\section{Partitioned matrices of maximal rank}\label{sec.partition}
Let the Hermitian matrix $H\in M_n(\C)$ be partitioned as follows:
\begin{equation}
	H = \begin{pmatrix}\label{eqn.defh}
		A & B \\
		B^\star & C
	\end{pmatrix},
\end{equation}
with $A\in M_k(\C)$, $1\le k\le n-1$. Let $V = K\oplus L$ be the decomposition of $V\simeq \C^n$ in accordance with the partition of $H$ in equation (\ref{eqn.defh}) so that $A$ is a map from $K$ to $K$, and $C$ is a map from $L$ to $L$. Let us further assume that $H$ is positive semi-definite. Let $k\in N(A)\subset K$, i.e. let $A k = 0$. For $v=(k,0)^T\in V$ we then have 
\begin{equation}
	v^\star H v=0.
\end{equation}
Since $H$ is positive semi-definite this implies (see \cite{matrixanalysis}) that we already have
\begin{equation}
	Hv = 0,
\end{equation}
or 
\begin{equation}\label{eqn.nulla}
	N(A) \subset N(B^\star).
\end{equation}
In a similar fashion we can establish that the null space of $C$ is contained in the null space of $B$:
\begin{equation}\label{eqn.nullc}
	N(C) \subset N(B).
\end{equation}
The relations for the null spaces are equivalent to these relations for the ranges:
\begin{align}
	R(B) & \subset R(A) \\
	R(B^\star) & \subset R(C).
\end{align}
Because of these properties $H$ is said to have the \emph{column inclusion property}\cite{matrixanalysis}. It follows from equations (\ref{eqn.nulla}) and (\ref{eqn.nullc}) that 
\begin{equation}
	N(A) \oplus N(C) \subset N(H).
\end{equation}
This implies that the rank of $H$ is less than or equal to the sum of the ranks of $A$ and $C$:
\begin{align}
	\rank\ H & = n - \dim N(H) \\
	& \le n - (\dim N(A) + \dim N(C)) \\
	& = k-\dim N(A) + (n-k) - \dim N(C) \\
	& = \rank\ A + \rank\ C
\end{align}
We have equality if and only if $N(H) = N(A)\oplus N(C)$. In the following, matrices $H$ for which this equality holds, will be of particular interest to us which is why we make the following definition:

\begin{definition}
	Let $H$ be a positive semi-definite Hermitian matrix that is partitioned as in equation (\ref{eqn.defh}). We say that $H$ is of \emph{maximal rank} if and only if $N(H) = N(A)\oplus N(C)$. 
\end{definition}

When $H$ is of maximal rank, it vanishes on
\begin{equation}
	N(A)\oplus N(C),
\end{equation}
and is positive definite when restricted to the sum of the ranges of $A$ and $C$:
\begin{equation}
	R(A) \oplus R(C).
\end{equation} 
We will use this property to extend results that are valid for positive definite matrices to matrices of maximal rank. Before we can state these results we need to introduce one more notion.

\section{The generalized determinant}\label{sec.gendet}
A Hermitian matrix $H$ defines a non-singular map from its range $R(H)$ to its range $R(H)$. If $H$ is singular its determinant vanishes. The determinant thus contains no information about $H$ restricted to $R(H)$. To recover this information we introduce a generalized determinant $\detp$:

\begin{definition}\label{def.gendet}
	Let $H$ be Hermitian. Let $\bar H$ be $H$ restricted to the range of $H$:
	\begin{equation}
		\bar H = H \vert_{R(H)} : R(H) \longrightarrow R(H) \\
	\end{equation}
	For $H\ne 0$ we then set 
	\begin{equation}
		\detp H = \det \bar H.
	\end{equation}
	For $H=0$ we set $\detp H = 1$.
\end{definition}

We note a number of properties of the generalized determinant:

\begin{lemma}
	Let $H$ be Hermitian of rank $r\le n$. Let $\lambda_i, i=1, \ldots, n$, be the eigenvalues of $H$. Let us assume that they are ordered in such a way that $\lambda_i = 0$, for $i>r$. We then have:
	\begin{enumerate}
		\item If $H$ is of full rank (i.e. if $r=n$) we have
		\begin{equation}
			\detp H = \det H = \prod_i \lambda_i
		\end{equation}
		\item For $r<n$ we have
		\begin{equation}
			\detp H = \prod_{i=1}^r \lambda_i
		\end{equation}
		\item We have
		\begin{equation}
			\detp H = \lim_{\epsilon \rightarrow 0} \frac{\det( H + \epsilon I )}{\epsilon^{n-r}}
		\end{equation}
		\item For $c>0$ we have
		\begin{equation}
			\detp c H = c^{r} \detp H
		\end{equation}
	\end{enumerate}   
\end{lemma}

\begin{proof}
	All of these identities follow from the fact that $H = U D U^\star$ for a unitary $U$ and a diagonal $D$ that contains the eigenvalues of $H$ on the diagonal. \qed
\end{proof}

We note that the generalized determinant does not have many properties the usual determinant has. In particular, it is not a continuous function of $H$, and the generalized determinant of the product of two matrices is not the product of the two determinants.

\section{The extensions}\label{sec.ext}
With the preparations in section \ref{sec.partition} and the definition of the generalized determinant we now want to extent three results: Fischer's inequality, the determinant equality for Schur's complement, and Babachiewicz's form of the inverse of a matrix.

\subsection{Fischer's inequality}
Fischer's inequality states that for a Hermitian positive semidefinite $H$ as in equation (\ref{eqn.defh}) we have 
\begin{equation}
	\det H \le \det A\; \det C.
\end{equation}
If $H$ is positive definite, we have equality if and only if $B=0$. If $H$ is singular, the determinant of $H$ vanishes and the inequality is no longer much of a constraint. We can also no longer infer that $B=0$ in the case of equality.

We can do better when $H$ is of maximal rank. In this case $H$ vanishes on $N(A)\oplus N(C)$ and is positive definite on 
\begin{equation}
	R(A)\oplus R(C).
\end{equation}
We now look at the restriction of $H$ and all its sub-matrices to this space and use the usual Fischer inequality there. As in definition \ref{def.gendet}, we denote the restriction of $H$ to its range by $\bar H$. We obtain
\begin{align}
	\detp H & = \det \bar H \\
	&\le \det \bar A\; \det \bar C \\
	& = \detp A\; \detp C.
\end{align}
Because we are looking at the restriction of $H$ to $R(A)\oplus R(C)$ and because it is positive definite on $R(A)\oplus R(C)$, we also get that $B$, when restricted to $R(C)$, vanishes if and only if equality holds above. Since $B$ vanishes on $N(C)$, this is the case if and only if
\begin{equation}
	B = 0.
\end{equation}
We thus have the following result:

\begin{theorem}
	Let a Hermitian $H$ be positive semidefinite and partitioned as in equation (\ref{eqn.defh}). If $H$ is of maximal rank then
	\begin{equation}
		\detp H \le \detp A\; \detp C,
	\end{equation} 
	with equality if and only if $B = 0$. 
\end{theorem}

\subsection{Schur complement}

\section{Matrix completion}

\section{Conclusion}

\begin{thebibliography}{MM}
	\bibitem{matrixanalysis} Roger A. Horn, Charles R. Johnson, \emph{Matrix Analysis}, 2nd Edition, Cambridge University Press, 2013.
	\bibitem{hayns} Emilie Virginia Haynsworth, Determination of the inertia of a partitioned Hermitian matrix,  Linear  Algebra and its Applications \textbf{1}, 1968, 73--81. 
	\bibitem{horn} Roger A. Horn, Basic  Properties  of  the  Schur  Complement. In Fuzhen Zhang (Ed.) The Schur complement and its application (17--46). Springer, 2005.
	\bibitem{rao} C. Radharkishna Rao, Linear Statistical Inference and its Applications, 2nd edition, John Wiley \& Sons, Inc, 2002.
	\bibitem{cottle} Richard W. Cottle, Manifestations of the Schur complement, Linear Algebra and its Applications \textbf{8}, 1974, 189--211. 
	\bibitem{smith} Smith, R. (2008). The positive definite completion problem revisited Linear Algebra and its Applications  429(7), 1442-1452. https://dx.doi.org/10.1016/j.laa.2008.04.020
	\bibitem{ouellette} Ouellette, D. (1981). Schur complements and statistics Linear Algebra and its Applications  36, 187-295. https://dx.doi.org/10.1016/0024-3795(81)90232-9
	\bibitem{roger} Penrose, R. (1955). A generalized inverse for matrices Mathematical Proceedings of the Cambridge Philosophical Society  51(3), 406-413. https://dx.doi.org/10.1017/s0305004100030401
	\bibitem{albert} Albert, A. (1969). Conditions for Positive and Nonnegative Definiteness in Terms of Pseudoinverses SIAM Journal on Applied Mathematics  17(2), 434-440. https://dx.doi.org/10.1137/0117041
\end{thebibliography}

\appendix

\section{Maybe ...}

\end{document}

### 2021-05-13 paper/paper_math.tex
Added the 2nd Theorem. Made a graphic.

@@
 \end{equation} 
-We will use this property to extend results that are valid for positive definite matrices to matrices of maximal rank. Before we can state these results we need to introduce one more notion.
+We will use this property to extend results that are valid for positive definite matrices to partitioned matrices of maximal rank.
+Before we can state these results we need to introduce one more notion.
 
@@
 
+Let us note that we arrived at this result in two steps.
+Because $H$ is positive semidefinite we know that $B$ restricted to $N(C)$ is zero.
+That the restriction of $B$ to the range $R(C)$ is also zero follows from Fischer's equality for the positive definite matrix that is $H$ restricted to $R(A)\oplus R(C)$.
+
 \subsection{Schur complement}
+We now turn our attention to the Schur complement. For a positive definite $H$ Schur showed that
+\begin{equation}
+	\det H = \det A\; \det H/A.
+\end{equation}
+We again focus our attention on $R(A)\oplus R(C)$, where $H$ is positive definite. Using the notation from the previous section we obtain:
+\begin{align}
+	\detp H & = \det \bar H \\
+	& = \det \bar A \; \det \bar H / \bar A \\
+	& = \detp A  \; \det \bar H / \bar A
+\end{align}
+We need to convince ourselves that the last determinant is equal to $\detp H/A$. To check this we need to show that 
+\begin{equation}
+	N(H/A) = N(C).
+\end{equation}
+We already know that $N(C) \subset N(H/A)$. To show the other inclusion let $l\in N(N/A)$ and set
+\begin{equation}
+	k = - A^+ B l.
+\end{equation}
+It then follows that 
+\begin{equation}
+	H\begin{pmatrix}
+		k\\
+		l
+	\end{pmatrix} = 0.
+\end{equation}
+Since $H$ is of maximal rank this implies that $(k,l)^T\in N(A)\oplus N(C)$. In particular, we have $w\in N(C)$. We thus obtain our second result:
+
+\begin{theorem} Let a Hermitian $H$ be positive semidefinite and partitioned as in equation (\ref{eqn.defh}). If $H$ is of maximal rank then
+	\begin{equation}
+		\detp H = \detp A\; \detp H/A.
+	\end{equation}	
+\end{theorem}
+
+\subsection{The inverse}
 
@@
 	\bibitem{matrixanalysis} Roger A. Horn, Charles R. Johnson, \emph{Matrix Analysis}, 2nd Edition, Cambridge University Press, 2013.
+	\bibitem{schur} J. Schur, Über Potenzreihen, die im Innern des Einheitskreises beschränkt sind, Journal für die reine und angewandte Mathematik \textbf{147}, 205 -- 232, 1917.
 	\bibitem{hayns} Emilie Virginia Haynsworth, Determination of the inertia of a partitioned Hermitian matrix,  Linear  Algebra and its Applications \textbf{1}, 1968, 73--81. 

### 2021-05-15 paper/paper_math.tex
Added the uniqueness results.

@@
 \section{The extensions}\label{sec.ext}
-With the preparations in section \ref{sec.partition} and the definition of the generalized determinant we now want to extent three results: Fischer's inequality, the determinant equality for Schur's complement, and Babachiewicz's form of the inverse of a matrix.
+With the preparations in section \ref{sec.partition} and the definition of the generalized determinant we now want to extent three results: Fischer's inequality, the determinant equality for Schur's complement, and Banachiewicz's form of the inverse of a matrix.
 
@@
 
-\begin{theorem}
+\begin{proposition}\label{prop.fischer}
 	Let a Hermitian $H$ be positive semidefinite and partitioned as in equation (\ref{eqn.defh}). If $H$ is of maximal rank then
@@
 	with equality if and only if $B = 0$. 
-\end{theorem}
+\end{proposition}
 
@@
 \end{equation}
-We again focus our attention on $R(A)\oplus R(C)$, where $H$ is positive definite. Using the notation from the previous section we obtain:
+We again focus our attention on $R(A)\oplus R(C)$ where $H$ is positive definite. Using the notation from the previous section we obtain:
 \begin{align}
@@
 We need to convince ourselves that the last determinant is equal to $\detp H/A$. To check this we need to show that 
-\begin{equation}
-	N(H/A) = N(C).
-\end{equation}
-We already know that $N(C) \subset N(H/A)$. To show the other inclusion let $l\in N(N/A)$ and set
+\begin{align}
+	N(H/A) & = N( C - B^\star A^+ B ) \\ 
+	& = N(C).
+\end{align}
+We already know that $N(C) \subset N(H/A)$. To show equality let $l\in N(N/A)$ and set
 \begin{equation}
@@
 
-\begin{theorem} Let a Hermitian $H$ be positive semidefinite and partitioned as in equation (\ref{eqn.defh}). If $H$ is of maximal rank then
+\begin{proposition}\label{prop.schur}
+ Let a Hermitian $H$ be positive semidefinite and partitioned as in equation (\ref{eqn.defh}). If $H$ is of maximal rank then
 	\begin{equation}
@@
 	\end{equation}	
+\end{proposition}
+
+\subsection{The inverse}
+The result concerns the inverse of $H$. When $H$ is positive definite its inverse can be written as
+\begin{equation}\label{eqn.bana}
+	H^{-1} = \begin{pmatrix}
+		A^{-1} + A^{-1}B(H/A)^{-1}B^\star A^{-1} & - A^{-1} B (H/A)^{-1} \\
+		- (H/A)^{-1} B^\star A^{-1} & (H/A)^{-1}
+	\end{pmatrix}.
+\end{equation} 
+This is a remarkable formula because to find the of $H$ we just need to invert $A$ and $H/A$.
+This formula was first established by Banachiewicz \cite{bana} (see also \cite[p.112]{frazer}).
+We now want to adapt this formula to our situation where $H$ might be singular but is of full rank.
+
+Again, we start by looking at the restriction $\bar H$ of $H$ to its range $R(A)\oplus R(C)$.
+$\bar H$ is positive definite and we can express its inverse in the form of equation (\ref{eqn.bana}).
+In the last section we established that the range and null spaces of $C$ and $H/A$ coincide so that by replacing the inverses with Moore-Penrose inverses we obtain the unique Moore-Penrose inverse of $H$.
+
+\begin{proposition}\label{prop.bana}
+	Let a Hermitian $H$ be positive semidefinite and partitioned as in equation (\ref{eqn.defh}). If $H$ is of maximal rank then its Moore-Penrose inverse is given by
+	\begin{equation}\label{eqn.banapenrose}
+	H^{+} = \begin{pmatrix}
+		A^{+} + A^{+}B(H/A)^{+}B^\star A^{+} & - A^{+} B (H/A)^{+} \\
+		- (H/A)^{+} B^\star A^{+} & (H/A)^{+}
+	\end{pmatrix}.
+	\end{equation}
+\end{proposition}
+
+This result can also be found in \cite{ouellette}.
+
+\section{Matrix completion}
+Let $H$ be a Hermitian matrix in $M_n(\C)$ that is partitioned as follows:
+\begin{equation}\label{eqn.defnewh}
+	H = \begin{pmatrix}
+		A & B & X \\
+		B^\star & C & D \\
+		X^\star & D^\star & E
+	\end{pmatrix}
+\end{equation}
+To denote sub-matrices of $H$ we use the notation from \cite{matrixanalysis}. For a set $\alpha\subset \{1, \ldots, n\}$ we let
+\begin{equation}
+	H[\alpha]
+\end{equation}
+be the sub-matrix of $H$ with row and column indices in $\alpha$. For $\alpha,\beta\subset \{1, \ldots, n\}$ we let
+\begin{equation}
+	H[\alpha, \beta] 
+\end{equation}
+be the sub-matrix with row indices in $\alpha$ and column indices in $\beta$. Now let $\alpha$ and $\beta$ be such that
+\begin{equation}
+	H[\alpha] = \begin{pmatrix}
+		A & B \\
+		B^\star & C
+	\end{pmatrix}
+\end{equation}
+and 
+\begin{equation}
+	H[\beta] = \begin{pmatrix}
+		C & D \\
+		D^\star & E
+	\end{pmatrix}.
+\end{equation}
+For $\gamma = \alpha\cap\beta$ we then have
+\begin{equation}
+	H[\gamma] = C,
+\end{equation}
+and for the matrix $X$ in the upper right corner we have
+\begin{equation}
+	H[ \alpha-\gamma, \beta-\gamma ] = X.	
+\end{equation}
+We now want to know under what conditions we can choose $X$ so that the matrix $H$ is positive semidefinite. Before we state the result we note this helpful theorem:
+
+\begin{theorem}\label{theo.albert}
+	Let $H$ be Hermitian and partitioned as in equation (\ref{eqn.defh}). Then these two statements are equivalent:
+	\begin{enumerate}
+		\item $H$ is positive semidefinite.
+		\item $A$ and $H/A$ are positive semidefinite, and $R(B)\subset R(A)$.
+	\end{enumerate}	
 \end{theorem}
 
-\subsection{The inverse}
+A proof of this result can be found in \cite{albert}. We now have:
 
-\section{Matrix completion}
+\begin{theorem}\label{theo.x}
+	Let $H\in M_n(\C)$ be partitioned as in (\ref{eqn.defnewh}) and let $H[\alpha]$ and $H[\beta]$ be positive semidefinite. Setting
+	\begin{align}
+		X & = B C^+ D \\
+		& = H[\alpha-\gamma, \gamma] H[\gamma]^+ H[\gamma, \beta-\gamma]
+	\end{align}
+	turns $H$ into a positive semidefinite matrix.
+\end{theorem}
 
+\begin{proof}
+	We apply theorem \ref{theo.albert} to the positive semidefinite matrices $H[\alpha]$ and $H[\beta]$ to obtain:
+	\begin{align}
+		C & \ge 0 \\
+		H[\alpha]/C & \ge 0 \label{eqn.hac}\\
+		H[\beta]/C  & \ge 0 \label{eqn.hbc} \\
+		R(B^\star) & \subset R(C) \label{eqn.bsc} \\				
+		R(D) & \subset R(C) \label{eqn.dsc}
+	\end{align}	
+	The Schur complement $H/C$ is given by
+	\begin{equation}
+		H/C = \begin{pmatrix}
+			A - B C^+ B^\star & X - B C^+ D \\
+			X^\star - D^\star C^+ B^\star & E - D^\star C^+ D
+		\end{pmatrix}.
+	\end{equation}
+	If we choose 
+	\begin{equation}
+		X = B C^+ D,
+	\end{equation}
+	and also recognize the expressions for $H[\alpha]/C$ and $H[\beta]/C$ we find
+	\begin{equation}
+		H/C = \begin{pmatrix}
+			H[\alpha]/C & 0 \\
+			0 & H[\beta]/C
+		\end{pmatrix}.
+	\end{equation}
+	Because of equations (\ref{eqn.hac}) and (\ref{eqn.hbc}) we have
+	\begin{equation}
+		H/C \ge 0.
+	\end{equation}
+	Equations (\ref{eqn.bsc}) and (\ref{eqn.dsc}) then ensure that
+	\begin{equation}
+		R(B^\star\; D ) \subset R(C).
+	\end{equation}
+	Since we also have $C\ge 0$ we can use the theorem \ref{theo.albert} one more time, this time in the other direction, to obtain
+	\begin{equation}
+		H \ge 0.
+	\end{equation} 
+	This completes the proof. \qed
+\end{proof}
+
+This result appeared already in \cite{smith}. We have provided a different proof that makes use of theorem \ref{theo.albert}.
+
+It turns out that the choice of $X$ in theorem \ref{theo.x} gives $H$ unique properties.
+It is in these characterizations of $X$ that we go beyond the results in \cite{smith} because we include the case in which $H$ is singular.
+The first result characterizes $X$ as the unique extension of $H$ that maximizes the determinant of $H$.
+If $H$ is non-singular we can just talk about the regular determinant of $H$.
+If $H$ is singular we have to use the generalized determinant that we introduced in section \ref{sec.gendet}.
+
+\begin{theorem}
+	Let $H\in M_n(\C)$ be partitioned as in (\ref{eqn.defnewh}) and let $H[\alpha]$ and $H[\beta]$ be positive semidefinite and of full rank. The choice 
+	\begin{equation}
+		X = B C^+ D
+	\end{equation}
+	is the unique choice for $X$ for which $H$ is positive semidefinite, of maximal rank and for which the (generalized) determinant is maximal.  
+\end{theorem}
+
+\begin{proof}
+We have shown in the last theorem that $H$ is positive semidefinite if we set $X = B C^+ D$. $H$ is also of maximal rank. In general we have
+\begin{equation}
+	\rank\ H \le \rank\ A + \rank\ C + \rank\ E.
+\end{equation}
+For a positive semidefinite matrix rank is additive over the Schur complement (see \cite{ouellette}) so that we actually have equality:
+\begin{align}
+	\rank\ H & = \rank\ C + \rank\ H/C \\
+	& = \rank\ C + \rank\ H[\alpha]/C + \rank\ H[\beta]/C \\
+	& = \rank\ C + \rank\ A + \rank\ E
+\end{align}
+To establish the last equality we have used the assumption that both $H[\alpha]$ and $H[\beta]$ are of full rank.
+Thus $X=B C^+ D$ turns $H$ into a matrix of full rank.
+We now want to show that it is the only such choice that also maximizes the determinant.
+
+Let us now assume that $H$ is positive semidefinite and of full rank so that we can make use of propositions \ref{prop.fischer} and \ref{prop.schur}. Because of proposition \ref{prop.schur} we have
+\begin{equation}
+	\detp H = \detp C \; \detp H/C.
+\end{equation}
+Because of proposition \ref{prop.fischer} we have
+\begin{equation}
+	\detp H/C \le \detp H[\alpha]/C + \detp H[\beta]/C,
+\end{equation}
+with equality if and only if
+\begin{equation}
+	X = B C^+ D.
+\end{equation}
+It follows that this $X$ is the unique choice that maximizes the determinant. \qed
+\end{proof}
+
+For a non-singular matrix $H$ we can use the determinant to find the inverse of $H$ (see \cite{matrixanalysis}):
+\begin{equation}\label{eqn.inverseformula}
+	H^{-1} = \frac{1}{\det H}\left(\frac{\partial}{\partial h_{ij}}\det H \right)^T
+\end{equation}
+Because we are expressing $X$ in terms of the given sub-matrices $A, \ldots, E$ of $H$, it is clear that the derivative in equation (\ref{eqn.inverseformula}) vanishes for indices $i$ and $j$ that denote elements of $X$ itself.
+$H^{-1}$ has zeroes in those places where the matrix $X$ sits in $H$.
+It turns out that this characteristic uniquely determines $X$, even if $H$ is only positive semidefinite and we have to talk about the Moore-Penrose inverse of $H$ instead.
+
+\begin{theorem}
+	Let $H\in M_n(\C)$ be partitioned as in (\ref{eqn.defnewh}) and let $H[\alpha]$ and $H[\beta]$ be positive semidefinite and of full rank. The choice 
+	\begin{equation}
+		X = B C^+ D
+	\end{equation}
+	is the unique choice for $X$ for which $H$ is of full rank and for which $H^{+}$ has zeroes in those places where $X$ sits in $H$.
+\end{theorem}
+
+\begin{proof}
+	Because $H$ is of full rank we can use proposition \ref{prop.bana} to express the Moore-Penrose inverse of $H$ in terms of $C^+$ and $(H/C)^+$.
+If $H^+$ has zeroes where $X$ is in $H$ then we must have
+	\begin{equation}
+		(H/C)^+ = \begin{pmatrix}
+			(H[\alpha]/C)^+ & 0 \\
+			0 & (H[\beta]/C)^+  
+		\end{pmatrix},
+	\end{equation}
+	which can only be the case if $X=B C^+ D$. \qed
+\end{proof}
+
+For completeness we give the Moore-Penrose inverse for $H$:
+
+\begin{equation}
+	H^+ = \begin{pmatrix}
+		(H[\alpha]/C)^+ & -(H[\alpha]/C)^+ B C^+ & 0 \\
+		- C^+ B^\star (H[\alpha]/C)^+ & \Xi & - C^+ D (H[\beta]/C)^+ \\
+		0 & - (H[\beta]/C)^+ D^\star C^+ & (H[\beta]/C)^+ 
+	\end{pmatrix},
+\end{equation}
+with
+\begin{equation}
+	\Xi = C^+ + C^+ B^\star (H[\alpha]/C)^+ B C^+ + C^+ D (H[\beta]/C)^+ D^\star C^+.
+\end{equation}
 \section{Conclusion}
@@
 	\bibitem{schur} J. Schur, Über Potenzreihen, die im Innern des Einheitskreises beschränkt sind, Journal für die reine und angewandte Mathematik \textbf{147}, 205 -- 232, 1917.
+	\bibitem{bana} T. Banachiewicz, Zur Berechnung der Determinanten, wie auch der Inversen, und zur darauf basierten Auflösung der Systeme linearer Gleichungen, Acta Astronom. S\'er. C 3, 41--67, 1937.
+	\bibitem{frazer} Frazer, R., Duncan, W., Collar, A. (1938). Elementary Matrices And Some Applications To Dynamics And Differential Equations, Cambridge University Press, 1938.
+	\bibitem{ouellette} Ouellette, D. (1981). Schur complements and statistics Linear Algebra and its Applications  36, 187-295. https://dx.doi.org/10.1016/0024-3795(81)90232-9
+	\bibitem{albert} Albert, A.
+(1969).
+Conditions for Positive and Nonnegative Definiteness in Terms of Pseudoinverses SIAM Journal on Applied Mathematics  17(2), 434-440.
+https://dx.doi.org/10.1137/0117041
+	\bibitem{smith} Smith, R. (2008). The positive definite completion problem revisited Linear Algebra and its Applications  429(7), 1442-1452. https://dx.doi.org/10.1016/j.laa.2008.04.020	
+	% older stuff
 	\bibitem{hayns} Emilie Virginia Haynsworth, Determination of the inertia of a partitioned Hermitian matrix,  Linear  Algebra and its Applications \textbf{1}, 1968, 73--81.
@@
 	\bibitem{cottle} Richard W. Cottle, Manifestations of the Schur complement, Linear Algebra and its Applications \textbf{8}, 1974, 189--211. 
-	\bibitem{smith} Smith, R. (2008). The positive definite completion problem revisited Linear Algebra and its Applications  429(7), 1442-1452. https://dx.doi.org/10.1016/j.laa.2008.04.020
-	\bibitem{ouellette} Ouellette, D. (1981). Schur complements and statistics Linear Algebra and its Applications  36, 187-295. https://dx.doi.org/10.1016/0024-3795(81)90232-9
 	\bibitem{roger} Penrose, R. (1955). A generalized inverse for matrices Mathematical Proceedings of the Cambridge Philosophical Society  51(3), 406-413. https://dx.doi.org/10.1017/s0305004100030401
-	\bibitem{albert} Albert, A.
-(1969).
-Conditions for Positive and Nonnegative Definiteness in Terms of Pseudoinverses SIAM Journal on Applied Mathematics  17(2), 434-440.
-https://dx.doi.org/10.1137/0117041
 \end{thebibliography}

### 2021-05-19 paper/paper_risk.tex
Added the Risk paper & added first chapter.

--- 2021-05-12 paper/paper_math.tex
@@
 \begin{document}
-\noindent{\Huge \textbf\textsf{{Matrix completion and semi-definite matrices.}}}
+\noindent{\Huge \textbf\textsf{{Completing correlation matrices}}}
 
@@
 
-\section{Partitioned matrices of maximal rank}\label{sec.partition}
-Let the Hermitian matrix $H\in M_n(\C)$ be partitioned as follows:
+\section{The completion criterium}
+Given an incomplete correlation matrix what is the right way to complete it? To motivate the answer we need to introduce some notation. Let 
 \begin{equation}
-	H = \begin{pmatrix}\label{eqn.defh}
+	X = ( x_1, \ldots, x_k )^T\in \R^k
+\end{equation}
+be a $k$ dimensional random variable with a density function given by a $k$ dimensional normal distribution
+\begin{align}
+	f(X) & = N_k[\mu, H_X]( X ) \\
+	& = (2\pi)^{-k/2}\det(H_X)^{-1/2}\exp\left(-\frac{1}{2}(X -\mu)^T H_X^{-1} (X-\mu)\right),
+\end{align}
+with mean $\mu$ and covariance matrix $H$.
+A crucial property of the normal distribution is that we condition on some of the $x_i$, $i=1, \dots, k$, we again obtain a normal distribution.
+Assume that we want to condition on the last $l$ components of $X$:
+\begin{equation}
+	Z = (x_{k-l+1}, \ldots, x_k)^T \in \R^l.
+\end{equation}
+Let us partition the matrix $H$ in a way that reflects this partition of $X$:
+\begin{equation}
+	H_X = \begin{pmatrix}
 		A & B \\
-		B^\star & C
+		B^T & C
 	\end{pmatrix},
 \end{equation}
-with $A\in M_k(\C)$, $1\le k\le n-1$.
-Let $V = K\oplus L$ be the decomposition of $V\simeq \C^n$ in accordance with the partition of $H$ in equation (\ref{eqn.defh}) so that $A$ is a map from $K$ to $K$, and $C$ is a map from $L$ to $L$.
-Let us further assume that $H$ is positive semi-definite.
-Let $k\in N(A)\subset K$, i.e.
-let $A k = 0$.
-For $v=(k,0)^T\in V$ we then have
+with $A\in M_{k-l}$, $C\in M_l$, and $B\in M_{k-l,l}$. Let $\bar X$ be the first $k-l$ components of $X$:
 \begin{equation}
-	v^\star H v=0.
+	\bar X = (x_1, \ldots, x_{k-l})^T\in \R^{k-l}
 \end{equation}
-Since $H$ is positive semi-definite this implies (see \cite{matrixanalysis}) that we already have
+Let us also assume that $X$ has zero mean:
 \begin{equation}
-	Hv = 0,
+	\mu = 0
 \end{equation}
-or 
-\begin{equation}\label{eqn.nulla}
-	N(A) \subset N(B^\star).
+The density function for $X$ conditioned on $Z$ is then given by
+\begin{equation}
+	f(X \vert Z ) = N_{k-l}[BC^{-1}Z, H_X/C ](\bar X),
 \end{equation}
-In a similar fashion we can establish that the null space of $C$ is contained in the null space of $B$:
-\begin{equation}\label{eqn.nullc}
-	N(C) \subset N(B).
+(see \cite[chapter 8]{rao} and \cite{cottle}) where $H_X/C$ is the Schur complement of $C$ in $H_X$:
+\begin{equation}\label{eqn.hxc}
+	H_X/C = A - B C^{-1}B^T
 \end{equation}
-The relations for the null spaces are equivalent to these relations for the ranges:
+(Emilie Haynsworth introduced the name and highlighted its usefulness in \cite{hayns}.
+Schur originally made use of it in \cite{schur}.
+For an overview of the properties of the Schur complement see \cite{horn}.) We see that we again obtain a normal distribution.
+We will see the Schur complement again soon.
+
+Now let us look at a second set of random variables 
+\begin{equation}
+	Y = (x_{k-l+1}, \ldots, x_k, x_{k+1}, \ldots, x_n)^T \in \R^{n-k},
+\end{equation}
+that has an overlap $Z$ with the random variables $X$. Let the density function for $Y$ be given by
+\begin{equation}
+	f(Y) = N_{n-k}[0,H_Y] ( Y),
+\end{equation}
+with
+\begin{equation}
+	H_Y = \begin{pmatrix}
+		C & D \\
+		D^T & E
+	\end{pmatrix}.
+\end{equation}
+Note that the submatrix $C$ is shared with $H_X$.
+In our context the matrix might have been obtained by a previous calibration of a model that is part of both $X$ and $Y$.
+If we were to condition on $Z$ we would find
+\begin{equation}
+	f(Y\vert Z) = N_{n-k-l}[ D^T C^{-1}Z,H_Y/C](\bar Y), 
+\end{equation}
+with
+\begin{equation}\label{eqn.hyc}
+	H_Y/C = E - D^TC^{-1}D
+\end{equation}
+and $\bar Y$ is the part of $Y$ that is not $Z$. If we want to describe $X$ and $Y$ together we are looking at the matrix
+\begin{equation}
+	H = \begin{pmatrix}
+		A & B & W \\
+		B^T & C & D \\
+		W^T & D^T & E
+	\end{pmatrix},
+\end{equation}
+with a yet to be determined matrix $W\in M_{k-l,n-l}$.
+Because $X$ and $Y$ share the random variables in $Z$ we can not make them independent.
+The next best thing that we can do is to demand that if we fix $Z$ the remaining parts of $X$ and $Y$ are independent.
+The combined density for $X$ and $Y$ when we condition on $Z$ is given by
+\begin{equation}
+	f( X,Y\vert Z) = N_{n-l}[ (B^T,D)^TC^{-1}Z, H/C](\bar X, \bar Y),
+\end{equation}
+with
 \begin{align}
-	R(B) & \subset R(A) \\
-	R(B^\star) & \subset R(C).
+	H/C & = \begin{pmatrix}
+		A - B C^{-1}B^T & W - B C^{-1} D \\
+		W^T - D^TC^{-1}B^T & E - D^TC^{-1}D
+	\end{pmatrix}\\
+  & = \begin{pmatrix}
+		H_X/C & W - B C^{-1} D \\
+		W^T - D^TC^{-1}B^T & H_Y/C
+	\end{pmatrix}.
 \end{align}
-Because of these properties $H$ is said to have the \emph{column inclusion property}\cite{matrixanalysis}. It follows from equations (\ref{eqn.nulla}) and (\ref{eqn.nullc}) that 
+We see that if we set
 \begin{equation}
-	N(A) \oplus N(C) \subset N(H).
+	W = BC^{-1}D
 \end{equation}
-This implies that the rank of $H$ is less than or equal to the sum of the ranks of $A$ and $C$:
-\begin{align}
-	\rank\ H & = n - \dim N(H) \\
-	& \le n - (\dim N(A) + \dim N(C)) \\
-	& = k-\dim N(A) + (n-k) - \dim N(C) \\
-	& = \rank\ A + \rank\ C
-\end{align}
-We have equality if and only if $N(H) = N(A)\oplus N(C)$.
-In the following, matrices $H$ for which this equality holds, will be of particular interest to us which is why we make the following definition:
-
-\begin{definition}
-	Let $H$ be a positive semi-definite Hermitian matrix that is partitioned as in equation (\ref{eqn.defh}). We say that $H$ is of \emph{maximal rank} if and only if $N(H) = N(A)\oplus N(C)$. 
-\end{definition}
-
-When $H$ is of maximal rank, it vanishes on
+the above matrix becomes block diagonal and the conditional densities obey
 \begin{equation}
-	N(A)\oplus N(C),
+	f(X,Y\vert Z ) = f(X\vert Z)f(Y\vert Z)
 \end{equation}
-and is positive definite when restricted to the sum of the ranges of $A$ and $C$:
-\begin{equation}
-	R(A) \oplus R(C).
-\end{equation} 
-We will use this property to extend results that are valid for positive definite matrices to matrices of maximal rank. Before we can state these results we need to introduce one more notion.
-
-\section{The generalized determinant}\label{sec.gendet}
-A Hermitian matrix $H$ defines a non-singular map from its range $R(H)$ to its range $R(H)$.
-If $H$ is singular its determinant vanishes.
-The determinant thus contains no information about $H$ restricted to $R(H)$.
-To recover this information we introduce a generalized determinant $\detp$:
-
-\begin{definition}\label{def.gendet}
-	Let $H$ be Hermitian. Let $\bar H$ be $H$ restricted to the range of $H$:
-	\begin{equation}
-		\bar H = H \vert_{R(H)} : R(H) \longrightarrow R(H) \\
-	\end{equation}
-	For $H\ne 0$ we then set 
-	\begin{equation}
-		\detp H = \det \bar H.
-	\end{equation}
-	For $H=0$ we set $\detp H = 1$.
-\end{definition}
-
-We note a number of properties of the generalized determinant:
-
-\begin{lemma}
-	Let $H$ be Hermitian of rank $r\le n$. Let $\lambda_i, i=1, \ldots, n$, be the eigenvalues of $H$. Let us assume that they are ordered in such a way that $\lambda_i = 0$, for $i>r$. We then have:
-	\begin{enumerate}
-		\item If $H$ is of full rank (i.e. if $r=n$) we have
-		\begin{equation}
-			\detp H = \det H = \prod_i \lambda_i
-		\end{equation}
-		\item For $r<n$ we have
-		\begin{equation}
-			\detp H = \prod_{i=1}^r \lambda_i
-		\end{equation}
-		\item We have
-		\begin{equation}
-			\detp H = \lim_{\epsilon \rightarrow 0} \frac{\det( H + \epsilon I )}{\epsilon^{n-r}}
-		\end{equation}
-		\item For $c>0$ we have
-		\begin{equation}
-			\detp c H = c^{r} \detp H
-		\end{equation}
-	\end{enumerate}   
-\end{lemma}
-
-\begin{proof}
-	All of these identities follow from the fact that $H = U D U^\star$ for a unitary $U$ and a diagonal $D$ that contains the eigenvalues of $H$ on the diagonal. \qed
-\end{proof}
-
-We note that the generalized determinant does not have many properties the usual determinant has.
-In particular, it is not a continuous function of $H$, and the generalized determinant of the product of two matrices is not the product of the two determinants.
-
-\section{The extensions}\label{sec.ext}
-With the preparations in section \ref{sec.partition} and the definition of the generalized determinant we now want to extent three results: Fischer's inequality, the determinant equality for Schur's complement, and Babachiewicz's form of the inverse of a matrix.
-
-\subsection{Fischer's inequality}
-Fischer's inequality states that for a Hermitian positive semidefinite $H$ as in equation (\ref{eqn.defh}) we have 
-\begin{equation}
-	\det H \le \det A\; \det C.
-\end{equation}
-If $H$ is positive definite, we have equality if and only if $B=0$.
-If $H$ is singular, the determinant of $H$ vanishes and the inequality is no longer much of a constraint.
-We can also no longer infer that $B=0$ in the case of equality.
-
-We can do better when $H$ is of maximal rank. In this case $H$ vanishes on $N(A)\oplus N(C)$ and is positive definite on 
-\begin{equation}
-	R(A)\oplus R(C).
-\end{equation}
-We now look at the restriction of $H$ and all its sub-matrices to this space and use the usual Fischer inequality there.
-As in definition \ref{def.gendet}, we denote the restriction of $H$ to its range by $\bar H$.
-We obtain
-\begin{align}
-	\detp H & = \det \bar H \\
-	&\le \det \bar A\; \det \bar C \\
-	& = \detp A\; \detp C.
-\end{align}
-Because we are looking at the restriction of $H$ to $R(A)\oplus R(C)$ and because it is positive definite on $R(A)\oplus R(C)$, we also get that $B$, when restricted to $R(C)$, vanishes if and only if equality holds above.
-Since $B$ vanishes on $N(C)$, this is the case if and only if
-\begin{equation}
-	B = 0.
-\end{equation}
-We thus have the following result:
-
-\begin{theorem}
-	Let a Hermitian $H$ be positive semidefinite and partitioned as in equation (\ref{eqn.defh}). If $H$ is of maximal rank then
-	\begin{equation}
-		\detp H \le \detp A\; \detp C,
-	\end{equation} 
-	with equality if and only if $B = 0$. 
-\end{theorem}
-
-\subsection{Schur complement}
-
-\section{Matrix completion}
+because the blocks in the diagonal are $H_X/C$ and $H_Y/C$ from equations (\ref{eqn.hxc}) and (\ref{eqn.hyc}).
+Given that $Z$ is part of both $X$ and $Y$ this is as much independence as we can ask for.
 
@@
 \begin{thebibliography}{MM}
-	\bibitem{matrixanalysis} Roger A. Horn, Charles R. Johnson, \emph{Matrix Analysis}, 2nd Edition, Cambridge University Press, 2013.
-	\bibitem{hayns} Emilie Virginia Haynsworth, Determination of the inertia of a partitioned Hermitian matrix,  Linear  Algebra and its Applications \textbf{1}, 1968, 73--81. 
-	\bibitem{horn} Roger A. Horn, Basic  Properties  of  the  Schur  Complement. In Fuzhen Zhang (Ed.) The Schur complement and its application (17--46). Springer, 2005.
 	\bibitem{rao} C. Radharkishna Rao, Linear Statistical Inference and its Applications, 2nd edition, John Wiley \& Sons, Inc, 2002.
 	\bibitem{cottle} Richard W. Cottle, Manifestations of the Schur complement, Linear Algebra and its Applications \textbf{8}, 1974, 189--211. 
-	\bibitem{smith} Smith, R. (2008). The positive definite completion problem revisited Linear Algebra and its Applications  429(7), 1442-1452. https://dx.doi.org/10.1016/j.laa.2008.04.020
+	\bibitem{hayns} Emilie Virginia Haynsworth, Determination of the inertia of a partitioned Hermitian matrix,  Linear  Algebra and its Applications \textbf{1}, 1968, 73--81.
+	\bibitem{schur} J. Schur, Über Potenzreihen, die im Innern des Einheitskreises beschränkt sind, Journal für die reine und angewandte Mathematik \textbf{147}, 205 -- 232, 1917.
+	\bibitem{horn} Roger A. Horn, Basic  Properties  of  the  Schur  Complement. In Fuzhen Zhang (Ed.) The Schur complement and its application (17--46). Springer, 2005.
+
+	% older stuff
+	\bibitem{matrixanalysis} Roger A. Horn, Charles R. Johnson, \emph{Matrix Analysis}, 2nd Edition, Cambridge University Press, 2013.
+	\bibitem{schur} J. Schur, Über Potenzreihen, die im Innern des Einheitskreises beschränkt sind, Journal für die reine und angewandte Mathematik \textbf{147}, 205 -- 232, 1917.
+	\bibitem{bana} T. Banachiewicz, Zur Berechnung der Determinanten, wie auch der Inversen, und zur darauf basierten Auflösung der Systeme linearer Gleichungen, Acta Astronom. S\'er. C 3, 41--67, 1937.
+	\bibitem{frazer} Frazer, R., Duncan, W., Collar, A. (1938). Elementary Matrices And Some Applications To Dynamics And Differential Equations, Cambridge University Press, 1938.
 	\bibitem{ouellette} Ouellette, D. (1981). Schur complements and statistics Linear Algebra and its Applications  36, 187-295. https://dx.doi.org/10.1016/0024-3795(81)90232-9
-	\bibitem{roger} Penrose, R. (1955). A generalized inverse for matrices Mathematical Proceedings of the Cambridge Philosophical Society  51(3), 406-413. https://dx.doi.org/10.1017/s0305004100030401
 	\bibitem{albert} Albert, A.
@@
 https://dx.doi.org/10.1137/0117041
+	\bibitem{smith} Smith, R. (2008). The positive definite completion problem revisited Linear Algebra and its Applications  429(7), 1442-1452. https://dx.doi.org/10.1016/j.laa.2008.04.020	
+	\bibitem{roger} Penrose, R. (1955). A generalized inverse for matrices Mathematical Proceedings of the Cambridge Philosophical Society  51(3), 406-413. https://dx.doi.org/10.1017/s0305004100030401
 \end{thebibliography}

### 2021-05-20 paper/paper_risk.tex
Started procedure. Added images.

@@
 with mean $\mu$ and covariance matrix $H$.
-A crucial property of the normal distribution is that we condition on some of the $x_i$, $i=1, \dots, k$, we again obtain a normal distribution.
+A crucial property of the normal distribution is that if we condition on some of the $x_i$, $i=1, \dots, k$, we again obtain a normal distribution.
 Assume that we want to condition on the last $l$ components of $X$:
@@
 \end{equation}
-Let us partition the matrix $H$ in a way that reflects this partition of $X$:
+We partition the matrix $H$ in a way that reflects this partition of $X$:
 \begin{equation}
@@
 \end{equation}
-Let us also assume that $X$ has zero mean:
-\begin{equation}
-	\mu = 0
-\end{equation}
-The density function for $X$ conditioned on $Z$ is then given by
+If we assume that $X$ has zero mean (i.e. $\mu=0$), the density function for $X$ conditioned on $Z$ is then given by
 \begin{equation}
@@
 \end{equation}
-and $\bar Y$ is the part of $Y$ that is not $Z$. If we want to describe $X$ and $Y$ together we are looking at the matrix
+and $\bar Y$ is the part of $Y$ that is not $Z$. If we now want to describe $X$ and $Y$ together we are looking at the matrix
 \begin{equation}
@@
 We see that if we set
-\begin{equation}
+\begin{equation}\label{eqn.choice}
 	W = BC^{-1}D
@@
 the above matrix becomes block diagonal and the conditional densities obey
-\begin{equation}
+\begin{equation}\label{eqn.independent}
 	f(X,Y\vert Z ) = f(X\vert Z)f(Y\vert Z)
@@
 Given that $Z$ is part of both $X$ and $Y$ this is as much independence as we can ask for.
+The way to choose the matrix $W$ and complete $H$ is to demand that $X$ and $Y$, when conditioned on their shared part $Z$, are independent.
+
+This choice of $W$ can be characterized in two more ways.
+Banachiewicz showed that the inverse of a matrix could be expressed using the Schur complement.
+Since the Schur complement $H/C$ is block diagonal the inverse of $H$ is given by
+\begin{equation}
+	H^{-1} = \begin{pmatrix}
+		(H_X/C)^{-1} & -(H_X/C)^{-1} B C^{-1} & 0 \\
+		- C^{-1} B^T (H_X/C)^{-1} & \Xi & - C^{-1} D (H_Y/C)^{-1} \\
+		0 & - (H_Y/C)^{-1} D^T C^{-1} & (H_Y/C)^{-1} 
+	\end{pmatrix},
+\end{equation}
+with
+\begin{equation}
+\Xi = C^{-1} + C^{-1} B^T (H_X/C)^{-1} B C^{-1} + C^{-1} D (H_Y/C)^{-1} D^T C^{-1}.
+\end{equation}
+We see that $H^{-1}$ has zeroes in the places where $W$ sits in $H$. The $W$ from equation (\ref{eqn.choice}) is the unique choice with that property.
+
+When looking at the determinant of $H$ we find another characterization of $W$. Using the Schur complement again, we find
+\begin{align}
+	\det H & = \det C \; \det H/C \\
+	& \le \det C \; \det H_X/C \; \det H_Y/C,\label{eqn.fischer}
+\end{align}
+where we have equality in equation (\ref{eqn.fischer}) if and only if the off-diagonal blocks vanish, i.e.
+if and only if $W$ is given by equation (\ref{eqn.choice}).
+This choice of $W$ is the unique choice that maximizes the determinant of $H$.
+
+The entropy $S$ of the normal distribution $N_n[\mu,H]$ is given by
+\begin{equation}
+	S = \frac{1}{2}\log \det H + \text{\ const.}
+\end{equation} 
+Because we are maximizing the determinant of $H$ we are also maximizing the entropy $S$ of the distribution described by $H$.
+The $H$ chosen by the $W$ from equation (\ref{eqn.choice}) is the one that puts the fewest constraints on the distribution of $X$ and $Y$.
+
+\section{The completion procedure}
+In the last section we have completed the matrix $H$ in the very simple case where there are two known matrices $H_X$ and $H_Y$ with just one shared part $C$.
+The general case is usually not this simple.
+Building on the result in \cite{grone} Smith \cite{smith} (see also \cite{ksd} for a more general treatment of the semidefinite case) showed how to solve the general problem by reducing it to a repeated application of the result from the previous section.
+We will discuss the required steps in this section by looking at a particular example, the cross currency model.
+
+Our model will consist of two interest rate models for the base currency $E$ and the foreign currency $A$.
+These two currencies are linked by an exchange rate $X$.
+We will model the two interest rates and the exchange rate with stochastic volatility models so that we have a total of six stochastic variables:
+\begin{equation}\label{eqn.stochvar}
+	E, \nu_E, A, \nu_A, X, \nu_X
+\end{equation}
+We will denote the corresponding correlation coefficients by 
+\begin{equation}
+	(a,b),
+\end{equation}
+where $a$ and $b$ are one of the stochastic variables from equation (\ref{eqn.stochvar}).
+In total there are 15 different correlation coefficients in the correlation matrix $H$ of our model.
+We assume that the coefficients
+\begin{equation}
+	(E,\nu_E)\text{\ and\ }(A,\nu_A)
+\end{equation}
+have already been determined in separate calibrations of the interest rate models.
+That leaves us with 13 free coefficients in $H$.
+This number of coefficients is too large for a stable calibration.
+Instead we will calibrate the coefficients
+\begin{equation}
+	(E,A), (E,X), (A,X),(X,\nu_X),
+\end{equation}
+and determine the remaining nine coefficients using the completion procedure (see figure \ref{fig.xccymodel}).
+
+\begin{figure}[hbt]
+  \begin{center}
+  	\includegraphics[width=10cm]{img/xccymodel.png}
+  \end{center}
+  \caption{The correlation matrix for the cross currency model.
+The coefficients $(x_E, \nu_E)$ and $(x_A, \nu_A)$ (the boxes with the solid lines) will be determined in two separate interest rate calibrations.
+The coefficients $(x_E, X_E^A)$, $(x_E, x_A)$, $(x_A, X_E^A)$, and $(X_E^A, \nu_{EA})$ (the boxes with the dashed lines) will be determined in the cross currency calibration itself.
+The rest will be determined using the completion procedure.}\label{fig.xccymodel}
+\end{figure}
+
+The first step to complete the matrix $H$ is to determine the graph corresponding to $H$.
+The vertices of the graph are given by the row (or column) indices of $H$.
+In our case these are identified with the stochastic variables in equation (\ref{eqn.stochvar}).
+We then obtain the graph for $H$ by connecting the vertices for which we have coefficients.
+I n our example we obtain the graph in figure \ref{fig.fxCalibrationGraph}.
+
+\begin{figure}[hbt]
+  \begin{center}
+  	\includegraphics[width=10cm]{img/fxCalibrationGraph.png}
+  \end{center}
+  \caption{The initial graph for the cross currency calibration.
+The coefficients $(E,\nu_E)$ and $(A,\nu_A)$ (dashed edges) are determined in separate interest rate calibrations.
+The other edges are determined in the cross currency calibration itself.}\label{fig.fxCalibrationGraph}
+\end{figure}
+
+Now that we have the graph we need to determine if the graph is chordal (see e.g.
+\cite{golumbic}).
+A graph is chordal if every loop of length four or more has a chord (i.e.
+an edge connecting two non-consecutive vertices in the loop).
+In our example this is easy to check.
+The largest loop is the central loop consisting of the vertices $E$, $A$, and $X$.
+This loop has just three elements and thus does not require a chord.
+The procedure that we are describing in this section requires a chordal graph.
+
+The next step is to identify the cliques in our graph.
+A clique is a maximal set of vertices that are all connected to each other (the subgraph induced by the set of vertices is complete).
+In our example there are four such cliques given by the central loop and the three arms of the graph:
+\begin{align}
+	\alpha_C & = \{E,A,X\}\\
+	\alpha_E & = \{E,\nu_E\}\\
+	\alpha_A & = \{A,\nu_A\}\\
+	\alpha_X & = \{X,\nu_X\}
+\end{align}  
+Note that the cliques have non-trivial intersections.
+We can use these intersections to construct a new graph.
+The vertices of this new graph are given by the cliques $\alpha_C, \ldots, \alpha_X$ and we connect two cliques if their intersection is not empty.
+
+\begin{figure}[hbt]
+  \begin{center}
+  	\includegraphics[width=10cm]{img/cliques.png}
+  \end{center}
+  \caption{a) The graph for the cross currency has four cliques.
+b) These cliques are the vertices of a new graph.
+An edge in this new graph connects two cliques that have a non-empty intersection.}\label{fig.cliques}
+\end{figure}
 
@@
 	\bibitem{horn} Roger A. Horn, Basic  Properties  of  the  Schur  Complement. In Fuzhen Zhang (Ed.) The Schur complement and its application (17--46). Springer, 2005.
+	\bibitem{grone} R. Grone, C.R. Johnson, E. Sa, H. Wolkowicz, Positive definite completion of partial Hermitian matrices, Linear Algebra Appl. 58, 109--124, 1984.
+	\bibitem{smith} R. L. Smith, The positive definite completion problem revisited, Linear Algebra and its Applications 429 (2008) 1442 -- 1452.
+	\bibitem{ksd} our other paper ...
+	\bibitem{golumbic} M. C. Golumbic, Algorithmic Graph Theory and Perfect Graphs, Elsevier, 2004.
 

### 2021-05-22 paper/paper_risk.tex
Finished the xccy example.

@@
 This number of coefficients is too large for a stable calibration.
-Instead we will calibrate the coefficients
+Instead we will only calibrate the coefficients
 \begin{equation}
@@
 \end{equation}
-and determine the remaining nine coefficients using the completion procedure (see figure \ref{fig.xccymodel}).
+and determine the remaining nine coefficients using the completion procedure described here (see figure \ref{fig.xccymodel}).
 
@@
   \caption{The correlation matrix for the cross currency model.
-The coefficients $(x_E, \nu_E)$ and $(x_A, \nu_A)$ (the boxes with the solid lines) will be determined in two separate interest rate calibrations.
-The coefficients $(x_E, X_E^A)$, $(x_E, x_A)$, $(x_A, X_E^A)$, and $(X_E^A, \nu_{EA})$ (the boxes with the dashed lines) will be determined in the cross currency calibration itself.
-The rest will be determined using the completion procedure.}\label{fig.xccymodel}
+The coefficients $(E, \nu_E)$ and $(A, \nu_A)$ (the boxes with the dashed lines) are determined in separate interest rate calibrations.
+The coefficients $(E, X)$, $(E, A)$, $(A, X)$, and $(X, \nu_X)$ (the boxes with the solid lines) will be determined in the cross currency calibration itself.
+The remaining coefficients are the result of the completion procedure.}\label{fig.xccymodel}
 \end{figure}
 
-The first step to complete the matrix $H$ is to determine the graph corresponding to $H$.
+The first step to complete the matrix $H$ is to determine the graph $\gamma$ corresponding to $H$.
 The vertices of the graph are given by the row (or column) indices of $H$.
@@
 We then obtain the graph for $H$ by connecting the vertices for which we have coefficients.
-I n our example we obtain the graph in figure \ref{fig.fxCalibrationGraph}.
+In our example we obtain the graph in figure \ref{fig.fxCalibrationGraph}.
 
@@
   \end{center}
-  \caption{The initial graph for the cross currency calibration.
+  \caption{The initial graph $\gamma$ for the cross currency calibration.
 The coefficients $(E,\nu_E)$ and $(A,\nu_A)$ (dashed edges) are determined in separate interest rate calibrations.
@@
 
-Now that we have the graph we need to determine if the graph is chordal (see e.g.
+Now that we have the graph we need to determine if the graph is chordal.
+A graph is chordal if every loop of length four or more has a chord (i.e.
+an edge connecting two non-consecutive vertices in the loop; see e.g.
 \cite{golumbic}).
-A graph is chordal if every loop of length four or more has a chord (i.e.
-an edge connecting two non-consecutive vertices in the loop).
 In our example this is easy to check.
@@
 The next step is to identify the cliques in our graph.
-A clique is a maximal set of vertices that are all connected to each other (the subgraph induced by the set of vertices is complete).
+A clique is a maximal set of vertices that are all connected to each other (the subgraph induced by the set of vertices is complete; again, see \cite{golumbic}).
 In our example there are four such cliques given by the central loop and the three arms of the graph:
@@
 Note that the cliques have non-trivial intersections.
-We can use these intersections to construct a new graph.
-The vertices of this new graph are given by the cliques $\alpha_C, \ldots, \alpha_X$ and we connect two cliques if their intersection is not empty.
+We can use these intersections to construct a new graph $\Gamma$.
+The vertices of this new graph are given by the cliques $\alpha_C, \ldots, \alpha_X$ and we connect two cliques if their intersection is not empty (see figure \ref{fig.cliques}).
 
@@
   \caption{a) The graph for the cross currency has four cliques.
-b) These cliques are the vertices of a new graph.
+b) These cliques are the vertices of a new graph $\Gamma$.
 An edge in this new graph connects two cliques that have a non-empty intersection.}\label{fig.cliques}
 \end{figure}
+
+Because the graph $gamma$ for the matrix $H$ is chordal, there exists a clique tree, i.e.
+a spanning tree for the graph $\Gamma$ that possesses the intersection property.
+A spanning tree for $\Gamma$ is a tree that has the same vertices as the graph itself.
+It possesses the intersection property if for any two vertices $\alpha_1$ and $\alpha_2$ the intersection $\alpha_1\cap\alpha_2$ is contained in all the vertices on the unique path in the tree that connects $\alpha_1$ to $\alpha_2$ (see \cite{blair} for the different characterizations of a chordal graph).
+In our example the graph $\Gamma$ is itself a tree.
+It is also easy to check that it possesses the intersection property.
+We will choose $\alpha_E = \{E, \nu_E\}$ as the root of the tree.
+
+\begin{figure}[hbt]
+  \begin{center}
+  	\includegraphics[width=10cm]{img/cliqueTree.png}
+  \end{center}
+  \caption{The clique tree for the graph $\gamma$.
+The numbers on the right denote the height in the tree if we choose $\alpha_E$ to be the root of the tree.
+The completion procedure proceeds from the top of the tree to the bottom.
+We complete the matrix by repeatedly using equation (\ref{eqn.choice}) from the previous section.}\label{fig.cliqueTree}
+\end{figure}
+
+We now start completing the matrix $H$ by repeatedly applying equation (\ref{eqn.choice}) from the previous section.
+We start with the cliques on the top of the tree and work our way down.
+We start with the clique $\alpha_X=\{X, \nu_X\}$.
+It intersects the clique $\alpha_C$ in the element $X$.
+This leads us to make the following identifications:
+\begin{align}
+	H_X & = \begin{pmatrix}
+		1 & (X, \nu_X) \\
+		& 1
+	\end{pmatrix} \\
+	H_Y & = \begin{pmatrix}
+		1 & (E,X) & (A,X) \\
+		& 1 & (E,A) \\
+		& & 1 
+	\end{pmatrix}
+\end{align}
+Note that the common matrix $C$ is just given by 1. Equation (\ref{eqn.choice}) now gives:
+\begin{align}
+	W & = ( (E,\nu_X), (A,\nu_X) ) \\
+	& = BC^{-1}D \\
+	& = (X, \nu_X) ( (E,X), (A,X) ) ) \\
+	& =  ( (X, \nu_X)(E,X), (X, \nu_X)(A,X) ) ), 
+\end{align}
+or 
+\begin{align}
+	(E,\nu_X) & = (X, \nu_X)(E,X) \\
+	(A,\nu_X) & = (X, \nu_X)(A,X).
+\end{align}
+Figure \ref{fig.firstStep} shows the first completion step in the graph $\gamma$.
+
+\begin{figure}[hbt]
+  \begin{center}
+  	\includegraphics[width=10cm]{img/firstStep.png}
+  \end{center}
+  \caption{We start the completion of the matrix $H$ with the cliques of height two.
+Applying equation (\ref{eqn.choice}) to the matrices given by the cliques $\alpha_X$ and $\alpha_C$ adds the dashed lines to the graph $\gamma$.}\label{fig.firstStep}
+\end{figure}
+
+We have one more clique with height two in the clique tree for $\gamma$.
+In the next step we take the clique $\alpha_A$ and combine it with the newly created clique from the last step.
+This time we make the identifications
+\begin{align}
+	H_X & = \begin{pmatrix}
+		1 & (A, \nu_A) \\
+		& 1
+	\end{pmatrix} \\
+	H_Y & = \begin{pmatrix}
+		1 & (A,E) & (A,X) & (A,\nu_X) \\
+		& 1 & (E,X) & (E,\nu_X) \\
+		& & 1 & (X,\nu_X) \\
+		& & & 1 
+	\end{pmatrix}
+\end{align}
+We apply equation (\ref{eqn.choice}) again to obtain:
+\begin{align}
+	W & = ( (E,\nu_A), (X,\nu_A), (\nu_X,\nu_A) ) \\
+	& = BC^{-1}D \\
+	& = (A,\nu_A) ( (E,A), (A,X), (A,\nu_X) ),
+\end{align}
+or
+\begin{align}
+	(E,\nu_A) & = (A,\nu_A) (E,A) \\
+	(X,\nu_A) & = (A,\nu_A)(A,X)\\
+	(\nu_X,\nu_A) & = (A,\nu_A) (A,\nu_X) \\
+	& = (A,\nu_A)(X, \nu_X)(A,X)
+\end{align}
+Figure \ref{fig.secondStep} shows the result of this step in the graph $\gamma$.
+
+\begin{figure}[hbt]
+  \begin{center}
+  	\includegraphics[width=10cm]{img/secondStep.png}
+  \end{center}
+  \caption{In the second step we combine the clique from the first step with the remaining clique of height two in the tree.
+This step adds the three edges $(\nu_X,\nu_A)$, $(X,\nu_A)$, and $(E,\nu_A)$.}\label{fig.secondStep}
+\end{figure}
+
+There are no more cliques of height two left.
+Looking at the graph in figure \ref{fig.secondStep} we see that only the edges connecting $\nu_E$ to the large clique that is the result of the last step are missing.
+These missing edges will be created in the last step.
+We make the following identifications:
+\begin{align}
+		H_X & = \begin{pmatrix}
+		1 & (E, \nu_E) \\
+		& 1
+	\end{pmatrix} \\
+	H_Y & = \begin{pmatrix}
+		1 & (E,A) & (E,\nu_A) & (E,X) & (E,\nu_X) \\
+		& 1 &  (A,\nu_A) & (A,X) & (A,\nu_X) \\
+		& & 1 & (\nu_A,X) & (\nu_A,\nu_X) \\
+		& & & 1 & (X,\nu_X) \\
+		& & & & 1 
+	\end{pmatrix}
+\end{align}
+Applying equation (\ref{eqn.choice}) one more time gives:
+\begin{align}
+	W &= ( (\nu_E,A),(\nu_E,\nu_A),(\nu_E,X),(\nu_E,\nu_X) ) \\ 
+	& = BC^{-1}D \\
+	& = (E, \nu_E) ( (E,A),(E,\nu_A),(E,X),(E,\nu_X) ), 
+\end{align}
+or
+\begin{align}
+	(\nu_E,A) & = (E, \nu_E) (E,A) \\
+	(\nu_E,\nu_A) & = (E, \nu_E) (E,\nu_A) \\
+	& = (E, \nu_E) (A,\nu_A) (E,A) \\
+	(\nu_E,X) & = (E, \nu_E) (E,X) \\
+	(\nu_E,\nu_X) & = (E, \nu_E) (E,\nu_X) \\
+	& = (E, \nu_E) (X, \nu_X)(E,X)
+\end{align}
+With this step we have added the last missing edges in the graph $\gamma$ and have obtained a positive definite completion of the matrix $H$.
 
@@
 	\bibitem{golumbic} M. C. Golumbic, Algorithmic Graph Theory and Perfect Graphs, Elsevier, 2004. 
+	\bibitem{blair} Jean R.
+S.
+Blair, Barry Peyton, An introduction to chordal graphs and clique trees, in Alan George, John R.
+Gilbert, Joseph W.
+H.
+Liu (Eds.), Graph Theory and Sparse Matrix Computation (pp.
+1 -- 29), The IMA Volumes in Mathematics and its Applications, 56, Springer Verlag, 1993.
 

### 2021-05-25 paper/paper_risk.tex
Added the case of several currencies.

@@
 
+\section{Several currencies}
+In the previous section we have just looked at one foreign currency $A$ with one exchange rate $X$.
+The procedure can easily be extended to an arbitrary number of foreign currencies.
+We will denote the foreign currencies by a capital letter:
+\begin{equation}
+	A, B, C, \ldots
+\end{equation}
+The exchange rate between the domestic currency $E$ and the foreign currency $A$ will now be denoted by 
+\begin{equation}
+	X_E^A.
+\end{equation}
+If $m_A$ is an amount in currency $A$ then $X_E^Am_A$ is the corresponding amount in currency $E$.
+If we are given the same correlation coefficients as in the previous section we obtain the graph in figure \ref{fig.several}.
+The corresponding graph of cliques is shown in figure \ref{fig.severalCliques}.
+
+\begin{figure}[hbt]
+  \begin{center}
+  	\includegraphics[width=10cm]{img/severalCurrencies.png}
+  \end{center}
+  \caption{The graph corresponding to the incomplete correlation matrix for five foreign currencies.}\label{fig.several}
+\end{figure}
+
+\begin{figure}[hbt]
+  \begin{center}
+  	\includegraphics[width=12cm]{img/severalCurrCliques.png}
+  \end{center}
+  \caption{The graph of cliques for the case of several foreign currencies.
+A clique tree is obtained by removing the lighter edges in the middle layer.
+It is again a tree of height two.}\label{fig.several}
+\end{figure}
+
 \section{Conclusion}
@@
 
-	% older stuff
-	\bibitem{matrixanalysis} Roger A. Horn, Charles R. Johnson, \emph{Matrix Analysis}, 2nd Edition, Cambridge University Press, 2013.
-	\bibitem{schur} J. Schur, Über Potenzreihen, die im Innern des Einheitskreises beschränkt sind, Journal für die reine und angewandte Mathematik \textbf{147}, 205 -- 232, 1917.
-	\bibitem{bana} T. Banachiewicz, Zur Berechnung der Determinanten, wie auch der Inversen, und zur darauf basierten Auflösung der Systeme linearer Gleichungen, Acta Astronom. S\'er. C 3, 41--67, 1937.
-	\bibitem{frazer} Frazer, R., Duncan, W., Collar, A. (1938). Elementary Matrices And Some Applications To Dynamics And Differential Equations, Cambridge University Press, 1938.
-	\bibitem{ouellette} Ouellette, D. (1981). Schur complements and statistics Linear Algebra and its Applications  36, 187-295. https://dx.doi.org/10.1016/0024-3795(81)90232-9
-	\bibitem{albert} Albert, A.
-(1969).
-Conditions for Positive and Nonnegative Definiteness in Terms of Pseudoinverses SIAM Journal on Applied Mathematics  17(2), 434-440.
-https://dx.doi.org/10.1137/0117041
-	\bibitem{smith} Smith, R. (2008). The positive definite completion problem revisited Linear Algebra and its Applications  429(7), 1442-1452. https://dx.doi.org/10.1016/j.laa.2008.04.020	
-	\bibitem{roger} Penrose, R. (1955). A generalized inverse for matrices Mathematical Proceedings of the Cambridge Philosophical Society  51(3), 406-413. https://dx.doi.org/10.1017/s0305004100030401
 \end{thebibliography}

### 2021-05-26 paper/paper_math.tex
Two papers, all results.

@@
 \end{equation}
-with $A\in M_k(\C)$, $1\le k\le n-1$.
+with $A\in M_k(\C)$, $C\in M_l(\C)$, $1\le k,l\le n-1$, $k+l=n$.
 Let $V = K\oplus L$ be the decomposition of $V\simeq \C^n$ in accordance with the partition of $H$ in equation (\ref{eqn.defh}) so that $A$ is a map from $K$ to $K$, and $C$ is a map from $L$ to $L$.
 Let us further assume that $H$ is positive semi-definite.
-Let $k\in N(A)\subset K$, i.e.
-let $A k = 0$.
-For $v=(k,0)^T\in V$ we then have
+Let $w\in N(A)\subset K$ be an element of the null space of $A$, i.e.
+let $A w = 0$.
+For $v=(W,0)^T\in V$ we then have
 \begin{equation}
@@
 \begin{equation}
-	Hv = 0,
+	Hv = \begin{pmatrix}
+		0 \\
+		B^\star w
+	\end{pmatrix} = 0,
 \end{equation}
@@
 \end{equation}
-The relations for the null spaces are equivalent to these relations for the ranges:
+The relations for the null spaces are equivalent to these relations for the ranges of $A$ and $C$:
 \begin{align}
@@
 
-We note that the generalized determinant does not have many properties the usual determinant has.
+We note that the generalized determinant does not have many of the properties the usual determinant has.
 In particular, it is not a continuous function of $H$, and the generalized determinant of the product of two matrices is not the product of the two determinants.
@@
 \end{align}
-We already know that $N(C) \subset N(H/A)$. To show equality let $l\in N(N/A)$ and set
+We already know that $N(C) \subset N(H/A)$. To show equality let $l\in N(H/A)$ and set
 \begin{equation}
@@
 \end{equation}
-Since $H$ is of maximal rank this implies that $(k,l)^T\in N(A)\oplus N(C)$. In particular, we have $w\in N(C)$. We thus obtain our second result:
+Since $H$ is of maximal rank this implies that $(k,l)^T\in N(A)\oplus N(C)$. In particular, we have $l\in N(C)$. We thus obtain our second result:
 

### 2021-05-26 paper/paper_risk.tex
Two papers, all results.

@@
 \section{The completion criterium}
-Given an incomplete correlation matrix what is the right way to complete it? To motivate the answer we need to introduce some notation. Let 
+Given a correlation matrix with unknown coefficients what is the right way to complete it? To motivate the answer we need to introduce some notation. Let 
 \begin{equation}
@@
 \end{align}
-with mean $\mu$ and covariance matrix $H$.
+with mean $\mu$ and covariance matrix $H_X$.
 A crucial property of the normal distribution is that if we condition on some of the $x_i$, $i=1, \dots, k$, we again obtain a normal distribution.
-Assume that we want to condition on the last $l$ components of $X$:
+Let us assume that we want to condition on the last $l$ components of $X$:
 \begin{equation}
@@
 For an overview of the properties of the Schur complement see \cite{horn}.) We see that we again obtain a normal distribution.
-We will see the Schur complement again soon.
+We will meet the Schur complement again soon.
 
@@
 Note that the submatrix $C$ is shared with $H_X$.
-In our context the matrix might have been obtained by a previous calibration of a model that is part of both $X$ and $Y$.
+In our context the matrix $C$ might have been obtained by a previous calibration of a model that is part of both $X$ and $Y$.
 If we were to condition on $Z$ we would find
@@
 
-This choice of $W$ can be characterized in two more ways.
-Banachiewicz showed that the inverse of a matrix could be expressed using the Schur complement.
+This choice of $W$ can be characterized in two more ways (see \cite{smith} and \cite{ksd}).
+Banachiewicz showed that the inverse of a matrix can be expressed using the Schur complement.
 Since the Schur complement $H/C$ is block diagonal the inverse of $H$ is given by
@@
 In the last section we have completed the matrix $H$ in the very simple case where there are two known matrices $H_X$ and $H_Y$ with just one shared part $C$.
-The general case is usually not this simple.
+The general case is usually more complicated.
 Building on the result in \cite{grone} Smith \cite{smith} (see also \cite{ksd} for a more general treatment of the semidefinite case) showed how to solve the general problem by reducing it to a repeated application of the result from the previous section.
-We will discuss the required steps in this section by looking at a particular example, the cross currency model.
+We will discuss the required steps in this section by looking at a particular example: the cross currency model.
 
-Our model will consist of two interest rate models for the base currency $E$ and the foreign currency $A$.
+Our model will consist of two interest rate models; one for the base currency $E$ and one the foreign currency $A$.
 These two currencies are linked by an exchange rate $X$.
@@
 We can use these intersections to construct a new graph $\Gamma$.
-The vertices of this new graph are given by the cliques $\alpha_C, \ldots, \alpha_X$ and we connect two cliques if their intersection is not empty (see figure \ref{fig.cliques}).
+The vertices of this new graph are given by the cliques $\alpha_C, \ldots, \alpha_X$ and we connect two cliques if their intersection is non-empty (see figure \ref{fig.cliques}).
 
@@
 
-Because the graph $gamma$ for the matrix $H$ is chordal, there exists a clique tree, i.e.
+Because the graph $\gamma$ for the matrix $H$ is chordal, there exists a clique tree, i.e.
 a spanning tree for the graph $\Gamma$ that possesses the intersection property.
@@
 It is also easy to check that it possesses the intersection property.
+In the next section we will see an example where the intersection property is not trivially fulfilled.
 We will choose $\alpha_E = \{E, \nu_E\}$ as the root of the tree.
@@
 The corresponding graph of cliques is shown in figure \ref{fig.severalCliques}.
+The corresponding clique tree is obtained by erasing the lines that connect the cliques
+\begin{equation}
+	\alpha_{CA}, \alpha_{CB}, \ldots, \alpha_{CG}
+\end{equation}
+on the middle layer. The intersection of any two of these cliques is given by
+\begin{equation}
+	\{E\} \subset \alpha_E.
+\end{equation}
+Since the path connecting the two cliques passes through $\alpha_E$ this tree satisfies the intersection property.
 
@@
 1 -- 29), The IMA Volumes in Mathematics and its Applications, 56, Springer Verlag, 1993.
-
 \end{thebibliography}
 
-\appendix
-
-\section{Maybe ...}
-
 \end{document}

### 2021-10-25 paper/paper_risk.tex
Added introduction, abstract, and conclusion.

@@
-\documentclass[12pt, oneside]{article} 
-\usepackage[a4paper]{geometry}              
-\usepackage{graphicx}
+\documentclass[11pt, a4paper]{article} 
+% \usepackage[a4paper]{geometry}              
 \usepackage{amsmath}
 \usepackage{amssymb}
-\usepackage[table]{xcolor}
-
-%SetFonts
-\usepackage[T1]{fontenc}
-\usepackage[bitstream-charter]{mathdesign}
-%SetFonts
-
-%define example environment 
-\newcounter{examplecounter}
-\newenvironment{example}%
-{%
-\small\begin{quote}%
-\refstepcounter{examplecounter}%
-\textbf{Example \arabic{examplecounter}}%
-\quad%
-}%
-{%\schluss%
-\end{quote}%
+\usepackage{mathrsfs}
+\usepackage{graphicx}
+\usepackage{hyperref}
+\hypersetup{
+    colorlinks=true,  % hyperlinks will be coloured
+    linkcolor=blue,   % hyperlink text will be blue
+    citecolor=blue
 }
-
-%define remark environment 
-\newcounter{remarkcounter}
-\newenvironment{remark}%
-{%
-\small\begin{quote}%
-\refstepcounter{remarkcounter}%
-\textbf{Remark \arabic{remarkcounter}}%
-\quad%
-}%
-{%\schluss%
-\end{quote}%
-}
-
-%define remark environment 
-\newcounter{questioncounter}
-\newenvironment{question}%
-{%
-\small\begin{quote}%
-\refstepcounter{questioncounter}%
-\textbf{Question \arabic{questioncounter}}%
-\quad%
-}%
-{%\schluss%
-\end{quote}%
-}
-
-%define important environment
-\newenvironment{important}{\begin{quote}%
-\textbf{Important:}%
-\quad
-}{%
-\end{quote}%
-}
-
-\newcommand{\qed}{\nobreak \ifvmode \relax \else
-      \ifdim\lastskip<1.5em \hskip-\lastskip
-      \hskip1.5em plus0em minus0.5em \fi \nobreak
-      \vrule height0.5em width0.5em depth0.25em\fi}
+\usepackage{xcolor}
+% \usepackage{a4wide}
 
@@
 
-\usepackage{lettrine}
+\begin{document}
+\title{Completing correlation matrices}
 
-\begin{document}
-\noindent{\Huge \textbf\textsf{{Completing correlation matrices}}}
+\author{Olaf Dreyer\thanks{OD Consulting}
+\and Horst K\"ohler \thanks{Commerzbank AG}
+\and Thomas Streuer \thanks{RIVACON GmbH} }
 
-\noindent \textit{by Horst Köhler, Thomas Streuer, Olaf Dreyer}
+\date{\today}
 
-\vspace{.5cm}
+\maketitle
+
+\begin{abstract}
+	We describe a way to complete a correlation matrix that is not fully specified.
+Such matrices arise in financial applications when the number of stochastic variables becomes large or when several smaller models are combined in a larger model.
+We argue that the proper completion to consider is the matrix that maximizes the entropy of the distribution described by the matrix.
+We then give a way to construct this matrix starting from the graph associated with the incomplete matrix.
+If this graph is chordal our construction will result in a proper correlation matrix.
+We demonstrate the construction using a cross-currency model with six stochastic variables.
+\end{abstract}
+
+\textbf{Keywords:} stochastic models, hybrid models, stochastic volatility, correlation matrices, graph theory, chordal graphs
+
+\textbf{JEL:} C02, C63, G12
 
 \section{Introduction}
+Modern financial models are driven by a large number of stochastic variables.
+These stochastic variables are usually correlated which creates the need to determine the coefficients of the underlying correlation matrix.
+Since the number of correlation coefficients increases quadratically with the number of stochastic variables, this number can become very large.
+While it would be desirable to calibrate all these coefficients this is in practice not possible because such a calibration procedure would be too slow, too unstable, or would require calibration instruments for which quotes can not be obtained.
+Hybrid models are another source of undetermined correlation coefficients.
+Two calibrated models might be parts of a larger model.
+While the coefficients for the individual models have been determined in separate calibration, the coefficients that mix variables from the two different models are undetermined.
+In a large model, with many smaller constituent models, the number of such coefficients can constitute the bulk of the matrix.
 
-\section{The completion criterium}
+Faced with this situation, it is tempting to set the missing coefficients to zero.
+The idea here is that any other choice seems to add arbitrary information.
+This approach has several shortcomings though.
+The most important of which is that one might end up with a matrix that is not positive semidefinite and thus not a valid correlation matrix.
+To obtain a valid matrix further steps are required.
+One possibility is to use the fact that the set of positive definite correlation matrices is an open set that contains the unit matrix.
+It is thus always possible to scale the off-diagonal elements of the matrix with a common scale factor $\lambda<1$ and obtain a positive definite matrix.
+Another possibility is to choose the nearest correlation matrix in a suitable topology (see e.g \cite{higham}).
+These additional steps might change all the coefficients of the matrix, including those coefficients, that have been obtained in previous calibrations.
+They also rely on somewhat arbitrary further assumptions like a particular notion of distance on the space of matrices.
+
+The approach presented here does not suffer from these shortcomings.
+We start with a particular criterium that singles out one particular completion.
+The criterium is distinguished by the fact that it puts the least amount of additional constraints on the distribution described by the completed matrix.
+As it turns out, this is not the matrix that contains zeroes in the places of unknown coefficients.
+Rather, it is the inverse of the matrix that contains zeroes in those places that were undetermined before.
+This criterium removes the arbitrary nature of other constructions.
+
+We then show how to construct this completion.
+The construction starts by looking at the graph associated with the incomplete matrix.
+The vertices of the graph are the rows (or columns) of the matrix.
+Since we are looking at a correlation matrix, the vertices correspond to the stochastic variables of the model.
+Two vertices are connected by an edge in the graph if the corresponding correlation coefficient in the matrix is know.
+It is here that we encounter the one condition that needs to be satisfied for the construction to work.
+The graph that we have just constructed needs to be chordal.
+A graph is chordal if every loop in the graph that has more than three edges contains a chord, i.e.
+an edge that connects two non-adjacent vertices in the loop.
+If this condition is satisfied we can follow a construction first introduced by \cite{grone} and refined by \cite{smith} to complete the matrix.
+Every step of the construction adds new vertices to the edge which in turn correspond to entries in the correlation matrix.
+The matrix obtained in this way will automatically be a positive-semidefinite so that no further steps need to be taken to obtain a valid correlation matrix.
+
+The paper now consists of three parts.
+In the next section we discuss the completion criterium.
+The matrix can be characterized in three different ways and we present them here.
+We then discuss the construction of the completion in section \ref{sec.procedure}.
+We do this by following the concrete example of a cross-currency model with two currencies and one exchange rate.
+The two models for the interest rate and the model for the exchange rate are assumed to be stochastic volatility models so that we end up with a total of six stochastic variables.
+We describe all the required steps using this model but the steps can easily be generalized to more complicated models.
+One such model is the $N$-currency model with $(N-1)$ foreign exchange rates that is described in section \ref{sec.several}.
+
+\section{The completion criterium}\label{sec.criterium}
 Given a correlation matrix with unknown coefficients what is the right way to complete it? To motivate the answer we need to introduce some notation. Let 
@@
 
-\section{The completion procedure}
+\section{The completion procedure}\label{sec.procedure}
 In the last section we have completed the matrix $H$ in the very simple case where there are two known matrices $H_X$ and $H_Y$ with just one shared part $C$.
@@
 
-\section{Several currencies}
+\section{Several currencies}\label{sec.several}
 In the previous section we have just looked at one foreign currency $A$ with one exchange rate $X$.
@@
 A clique tree is obtained by removing the lighter edges in the middle layer.
-It is again a tree of height two.}\label{fig.several}
+It is again a tree of height two.}\label{fig.severalCliques}
 \end{figure}
 
-\section{Conclusion}
+\section{Conclusion}\label{sec.conclusion}
+In this paper we have described a method to complete an insufficiently specified correlation matrix. Because financial models are getting more complex such matrices are a common occurrence.
+
+\section*{Acknowledgement}
+The authors would like to thank Patrick B\"uchel for his support during the creation of this work; and Nataliya Koval, and Roland Seydel for insightful discussions on the subject.
 
 \begin{thebibliography}{MM}
-	\bibitem{rao} C. Radharkishna Rao, Linear Statistical Inference and its Applications, 2nd edition, John Wiley \& Sons, Inc, 2002.
-	\bibitem{cottle} Richard W. Cottle, Manifestations of the Schur complement, Linear Algebra and its Applications \textbf{8}, 1974, 189--211. 
-	\bibitem{hayns} Emilie Virginia Haynsworth, Determination of the inertia of a partitioned Hermitian matrix,  Linear  Algebra and its Applications \textbf{1}, 1968, 73--81.
-	\bibitem{schur} J. Schur, Über Potenzreihen, die im Innern des Einheitskreises beschränkt sind, Journal für die reine und angewandte Mathematik \textbf{147}, 205 -- 232, 1917.
-	\bibitem{horn} Roger A. Horn, Basic  Properties  of  the  Schur  Complement. In Fuzhen Zhang (Ed.) The Schur complement and its application (17--46). Springer, 2005.
-	\bibitem{grone} R. Grone, C.R. Johnson, E. Sa, H. Wolkowicz, Positive definite completion of partial Hermitian matrices, Linear Algebra Appl. 58, 109--124, 1984.
-	\bibitem{smith} R. L. Smith, The positive definite completion problem revisited, Linear Algebra and its Applications 429 (2008) 1442 -- 1452.
-	\bibitem{ksd} our other paper ...
-	\bibitem{golumbic} M. C. Golumbic, Algorithmic Graph Theory and Perfect Graphs, Elsevier, 2004. 
-	\bibitem{blair} Jean R.
-S.
-Blair, Barry Peyton, An introduction to chordal graphs and clique trees, in Alan George, John R.
-Gilbert, Joseph W.
+	\bibitem[Higham, 2002]{higham}, Higham, N. J. (2002). Computing the nearest correlation matrix - a problem from finance. \textit{IMA Journal of Numerical Analysis} 22: 329 - 343.
+		\bibitem[Grone et~al., 1984]{grone} Grone R., Johnson C.R., Sa E., Wolkowicz H. (1984). Positive definite completion of partial Hermitian matrices. \textit{Linear Algebra Appl.} 58: 109--124.
+	\bibitem[Smith, 2008]{smith} Smith, R. L. (2008). The positive definite completion problem revisited. \textit{Linear Algebra and its Applications} 429:1442 -- 1452.
+	\bibitem[Rao, 2002]{rao} C. R. Rao (2002). \textit{Linear Statistical Inference and its Applications, 2nd edition}. John Wiley \& Sons, Inc.
+	\bibitem[Cottle, 1974]{cottle} Cottle, R. W. (1974). Manifestations of the Schur complement. \textit{Linear Algebra and its Applications} 8:189--211. 
+	\bibitem[Haynsworth, 1968]{hayns} Haynsworth, E.V. (1968), Determination of the inertia of a partitioned Hermitian matrix,  \textit{Linear  Algebra and its Applications} 1:73--81.
+	\bibitem[Schur, 1917]{schur} Schur, J. (1917). \"Uber Potenzreihen, die im Innern des Einheitskreises beschr\"ankt sind. \textit{Journal f\"ur die reine und angewandte Mathematik} 147: 205--232.
+	\bibitem[Horn, 2005]{horn} Horn, R. A. (2005). Basic  Properties  of  the  Schur  Complement, 17--46. In Fuzhen Zhang (Ed.) \textit{The Schur complement and its application}. Springer.
+	\bibitem[Dreyer et~al., 2021] {ksd} our other paper ...
+	\bibitem[Golumbic, 2004]{golumbic} Golumbic, M. C. (2004). \textit{Algorithmic Graph Theory and Perfect Graphs}. Elsevier. 
+	\bibitem[Blair et~al., 1993]{blair} Blair, J.
+R.
+S., Peyton, B.
+(1993).
+An introduction to chordal graphs and clique trees, 1 -- 29.
+In George, A., Gilbert, J.
+R., Liu, J.
+W.
 H.
-Liu (Eds.), Graph Theory and Sparse Matrix Computation (pp.
-1 -- 29), The IMA Volumes in Mathematics and its Applications, 56, Springer Verlag, 1993.
+(Eds.).
+\textit{Graph Theory and Sparse Matrix Computation}, The IMA Volumes in Mathematics and its Applications 56.
+Springer.
 \end{thebibliography}

### 2021-11-16 paper/paper_math.tex
thomas's notes on risk paper; formatting of math paper

@@
-\documentclass[12pt, oneside]{article} 
-\usepackage[a4paper]{geometry}              
-\usepackage{graphicx}
+\documentclass[11pt, a4paper]{article} 
+% \usepackage[a4paper]{geometry}              
 \usepackage{amsmath}
 \usepackage{amssymb}
-\usepackage[table]{xcolor}
-
-%SetFonts
-\usepackage[T1]{fontenc}
-\usepackage[bitstream-charter]{mathdesign}
-%SetFonts
-
-%define example environment 
-\newcounter{examplecounter}
-\newenvironment{example}%
-{%
-\small\begin{quote}%
-\refstepcounter{examplecounter}%
-\textbf{Example \arabic{examplecounter}}%
-\quad%
-}%
-{%\schluss%
-\end{quote}%
+\usepackage{amsthm}
+\usepackage{mathrsfs}
+\usepackage{graphicx}
+\usepackage{hyperref}
+\hypersetup{
+    colorlinks=true,  % hyperlinks will be coloured
+    linkcolor=blue,   % hyperlink text will be blue
+    citecolor=blue
 }
-
-%define remark environment 
-\newcounter{remarkcounter}
-\newenvironment{remark}%
-{%
-\small\begin{quote}%
-\refstepcounter{remarkcounter}%
-\textbf{Remark \arabic{remarkcounter}}%
-\quad%
-}%
-{%\schluss%
-\end{quote}%
-}
-
-%define remark environment 
-\newcounter{questioncounter}
-\newenvironment{question}%
-{%
-\small\begin{quote}%
-\refstepcounter{questioncounter}%
-\textbf{Question \arabic{questioncounter}}%
-\quad%
-}%
-{%\schluss%
-\end{quote}%
-}
-
-%define important environment
-\newenvironment{important}{\begin{quote}%
-\textbf{Important:}%
-\quad
-}{%
-\end{quote}%
-}
-
-\newcommand{\qed}{\nobreak \ifvmode \relax \else
-      \ifdim\lastskip<1.5em \hskip-\lastskip
-      \hskip1.5em plus0em minus0.5em \fi \nobreak
-      \vrule height0.5em width0.5em depth0.25em\fi}
+\usepackage{xcolor}
+% \usepackage{a4wide}
 
@@
 \newtheorem{definition}{Definition}
-\newenvironment{proof}[1][Proof]{\begin{trivlist}
-\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
+%\newenvironment{proof}[1][Proof]{\begin{trivlist}
+%\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
 
@@
 
-\usepackage{lettrine}
+\begin{document}
+\title{Matrix completion and semidefinite matrics}
 
-\begin{document}
-\noindent{\Huge \textbf\textsf{{Matrix completion and semi-definite matrices.}}}
+\author{Olaf Dreyer\thanks{OD Consulting}
+\and Horst K\"ohler \thanks{Commerzbank AG}
+\and Thomas Streuer \thanks{RIVACON GmbH} }
 
-\noindent \textit{by Horst Köhler, Thomas Streuer, Olaf Dreyer}
+\date{\today}
 
-\vspace{.5cm}
+\maketitle
+
+\begin{abstract} Here comes the abstract.
+\end{abstract}
+
+\emph{AMS classification:} 15A15; 15A18; 15A83; 15B48
+ \\
+
+\emph{Keywords:} positive definite; positive semidefinite; matrix completions;
 
 \section{Introduction}
+And some introduction.
+
+The Schur complement needs to be introduced here.
 
@@
 Let $V = K\oplus L$ be the decomposition of $V\simeq \C^n$ in accordance with the partition of $H$ in equation (\ref{eqn.defh}) so that $A$ is a map from $K$ to $K$, and $C$ is a map from $L$ to $L$.
-Let us further assume that $H$ is positive semi-definite.
-Let $w\in N(A)\subset K$ be an element of the null space of $A$, i.e.
+Let us further assume that $H$ is positive semidefinite.
+Let $w\in N(A)\subset K$ be an element of the nullspace of $A$, i.e.
 let $A w = 0$.
-For $v=(W,0)^T\in V$ we then have
+For $v=(w,0)^T\in V$ we then have
 \begin{equation}
@@
 \end{equation}
-Since $H$ is positive semi-definite this implies (see \cite{matrixanalysis}) that we already have
+Since $H$ is positive semidefinite this implies (see \cite{matrixanalysis}) that we already have
 \begin{equation}
@@
 \begin{definition}
-	Let $H$ be a positive semi-definite Hermitian matrix that is partitioned as in equation (\ref{eqn.defh}). We say that $H$ is of \emph{maximal rank} if and only if $N(H) = N(A)\oplus N(C)$. 
+	Let $H$ be a positive semidefinite Hermitian matrix that is partitioned as in equation (\ref{eqn.defh}). We say that $H$ is of \emph{maximal rank} if and only if $N(H) = N(A)\oplus N(C)$. 
 \end{definition}
@@
 \begin{proof}
-	All of these identities follow from the fact that $H = U D U^\star$ for a unitary $U$ and a diagonal $D$ that contains the eigenvalues of $H$ on the diagonal. \qed
+	All of these identities follow from the fact that $H = U D U^\star$ for a unitary $U$ and a diagonal $D$ that contains the eigenvalues of $H$ on the diagonal.
 \end{proof}
@@
 We note that the generalized determinant does not have many of the properties the usual determinant has.
-In particular, it is not a continuous function of $H$, and the generalized determinant of the product of two matrices is not the product of the two determinants.
+In particular, it is not a continuous function of $H$, and the generalized determinant of the product of two matrices is not the product of the two generalized determinants.
 
 \section{The extensions}\label{sec.ext}
-With the preparations in section \ref{sec.partition} and the definition of the generalized determinant we now want to extent three results: Fischer's inequality, the determinant equality for Schur's complement, and Banachiewicz's form of the inverse of a matrix.
+With the preparations in section \ref{sec.partition}, and the definition of the generalized determinant in the last section, we now want to extent three results: Fischer's inequality, the determinant equality for Schur's complement, and Banachiewicz's form of the inverse of a matrix.
 
 \subsection{Fischer's inequality}
-Fischer's inequality states that for a Hermitian positive semidefinite $H$ as in equation (\ref{eqn.defh}) we have 
+Fischer's inequality states that for a Hermitian positive semidefinite $H$ as in equation (\ref{eqn.defh}) we have (see e.g. \cite{horn}):
 \begin{equation}
@@
 \end{equation}
-If $H$ is positive definite, we have equality if and only if $B=0$.
+Furthermore, if $H$ is positive definite, we have equality if and only if $B=0$.
 If $H$ is singular, the determinant of $H$ vanishes and the inequality is no longer much of a constraint.
@@
 This formula was first established by Banachiewicz \cite{bana} (see also \cite[p.112]{frazer}).
-We now want to adapt this formula to our situation where $H$ might be singular but is of full rank.
+We now want to adapt this formula to our situation where $H$ might be singular but is of maximal rank.
 
@@
 	\end{equation} 
-	This completes the proof. \qed
+	This completes the proof.
 \end{proof}
@@
 \begin{theorem}
-	Let $H\in M_n(\C)$ be partitioned as in (\ref{eqn.defnewh}) and let $H[\alpha]$ and $H[\beta]$ be positive semidefinite and of full rank. The choice 
+	Let $H\in M_n(\C)$ be partitioned as in (\ref{eqn.defnewh}) and let $H[\alpha]$ and $H[\beta]$ be positive semidefinite and of maximal rank. The choice 
 	\begin{equation}
@@
 \end{align}
-To establish the last equality we have used the assumption that both $H[\alpha]$ and $H[\beta]$ are of full rank.
-Thus $X=B C^+ D$ turns $H$ into a matrix of full rank.
+To establish the last equality we have used the assumption that both $H[\alpha]$ and $H[\beta]$ are of maximal rank.
+Thus $X=B C^+ D$ turns $H$ into a matrix of maximal rank.
 We now want to show that it is the only such choice that also maximizes the determinant.
 
-Let us now assume that $H$ is positive semidefinite and of full rank so that we can make use of propositions \ref{prop.fischer} and \ref{prop.schur}. Because of proposition \ref{prop.schur} we have
+Let us now assume that $H$ is positive semidefinite and of maximal rank so that we can make use of propositions \ref{prop.fischer} and \ref{prop.schur}. Because of proposition \ref{prop.schur} we have
 \begin{equation}
@@
 \end{equation}
-It follows that this $X$ is the unique choice that maximizes the determinant. \qed
+It follows that this $X$ is the unique choice that maximizes the determinant.
 \end{proof}
@@
 	\end{equation}
-	is the unique choice for $X$ for which $H$ is of full rank and for which $H^{+}$ has zeroes in those places where $X$ sits in $H$.
+	is the unique choice for $X$ for which $H$ is of maximal rank and for which $H^{+}$ has zeroes in those places where $X$ sits in $H$.
 \end{theorem}
@@
 \begin{proof}
-	Because $H$ is of full rank we can use proposition \ref{prop.bana} to express the Moore-Penrose inverse of $H$ in terms of $C^+$ and $(H/C)^+$.
+	Because $H$ is of maximal rank we can use proposition \ref{prop.bana} to express the Moore-Penrose inverse of $H$ in terms of $C^+$ and $(H/C)^+$.
 If $H^+$ has zeroes where $X$ is in $H$ then we must have
@@
 	\end{equation}
-	which can only be the case if $X=B C^+ D$. \qed
+	which can only be the case if $X=B C^+ D$.
 \end{proof}

### 2021-11-16 paper/paper_risk.tex
thomas's notes on risk paper; formatting of math paper

@@
 While it would be desirable to calibrate all these coefficients this is in practice not possible because such a calibration procedure would be too slow, too unstable, or would require calibration instruments for which quotes can not be obtained.
+The issue of stability is especially important in a risk neutral setup.
+Daily calibrations might display undesired fluctuations if the calibration is not stable.
 Hybrid models are another source of undetermined correlation coefficients.
@@
 We then discuss the construction of the completion in section \ref{sec.procedure}.
-We do this by following the concrete example of a cross-currency model with two currencies and one exchange rate.
-The two models for the interest rate and the model for the exchange rate are assumed to be stochastic volatility models so that we end up with a total of six stochastic variables.
-We describe all the required steps using this model but the steps can easily be generalized to more complicated models.
-One such model is the $N$-currency model with $(N-1)$ foreign exchange rates that is described in section \ref{sec.several}.
+We do this by following the concrete example of a cross-currency model consisting of two interest rate models that are linked by one exchange rate.
+The two interest rate models and the model for the exchange rate are assumed to be stochastic volatility models so that we end up with a total of six stochastic variables.
+We will use this model to describe the all the required steps.
+These steps can easily be generalized to more complicated models.
+The $N$-currency model with $(N-1)$ foreign exchange rates is one such model.
+It is described in section \ref{sec.several}.
 
@@
 \begin{equation}
-	f(X \vert Z ) = N_{k-l}[BC^{-1}Z, H_X/C ](\bar X),
+	f(\bar X \vert Z ) = N_{k-l}[BC^{-1}Z, H_X/C ](\bar X),
 \end{equation}
@@
 \begin{equation}
-	Y = (x_{k-l+1}, \ldots, x_k, x_{k+1}, \ldots, x_n)^T \in \R^{n-k},
+	Y = (x_{k-l+1}, \ldots, x_k, x_{k+1}, \ldots, x_n)^T \in \R^{n-k+l},
 \end{equation}
@@
 \begin{equation}
-	f(Y) = N_{n-k}[0,H_Y] ( Y),
+	f(Y) = N_{n-k+l}[0,H_Y] ( Y),
 \end{equation}
@@
 \begin{equation}
-	f(Y\vert Z) = N_{n-k-l}[ D^T C^{-1}Z,H_Y/C](\bar Y), 
+	f(\bar Y\vert Z) = N_{n-k-l}[ D^T C^{-1}Z,H_Y/C](\bar Y), 
 \end{equation}
@@
 \begin{equation}
-	f( X,Y\vert Z) = N_{n-l}[ (B^T,D)^TC^{-1}Z, H/C](\bar X, \bar Y),
+	f( \bar X,\bar Y\vert Z) = N_{n-l}[ (B^T,D)^TC^{-1}Z, H/C](\bar X, \bar Y),
 \end{equation}
@@
 \begin{equation}\label{eqn.independent}
-	f(X,Y\vert Z ) = f(X\vert Z)f(Y\vert Z)
+	f(\bar X,\bar Y\vert Z ) = f(\bar X\vert Z)f(\bar Y\vert Z)
 \end{equation}
@@
 
-Our model will consist of two interest rate models; one for the base currency $E$ and one the foreign currency $A$.
+Our model will consist of two interest rate models; one for the domestic currency $E$ and one the foreign currency $A$.
 These two currencies are linked by an exchange rate $X$.

### 2021-11-19 paper/paper_risk.tex
final additions

@@
 
-\author{Olaf Dreyer\thanks{OD Consulting}
+\author{Olaf Dreyer\thanks{OD Consulting, email: olaf.dreyer@gmail.com}
 \and Horst K\"ohler \thanks{Commerzbank AG}
@@
 Two calibrated models might be parts of a larger model.
-While the coefficients for the individual models have been determined in separate calibration, the coefficients that mix variables from the two different models are undetermined.
+While the coefficients for the individual models have been determined in separate calibrations, the coefficients that mix variables from the two different models are undetermined.
 In a large model, with many smaller constituent models, the number of such coefficients can constitute the bulk of the matrix.
@@
 The general case is usually more complicated.
-Building on the result in \cite{grone} Smith \cite{smith} (see also \cite{ksd} for a more general treatment of the semidefinite case) showed how to solve the general problem by reducing it to a repeated application of the result from the previous section.
+Building on the result in \cite{grone} Smith \cite{smith} (see also \cite{georgescu}) showed how to solve the general problem by reducing it to a repeated application of the result from the previous section.
 We will discuss the required steps in this section by looking at a particular example: the cross currency model.
@@
 \section{Conclusion}\label{sec.conclusion}
-In this paper we have described a method to complete an insufficiently specified correlation matrix. Because financial models are getting more complex such matrices are a common occurrence.
+The increasing size of financial models has made insufficiently specified correlation matrices a common occurrence.
+The larger the model the fewer correlation coefficients are usually known.
+We thus need a rational for how to deal with these incomplete matrices as well as a robust method for implementing this rational.
+In this paper we have supplied both.
+We have argued that the correct way to complete the matrix is the one that maximizes the entropy of the distribution described by the matrix.
+This is the matrix with the maximal determinant.
+We have then shown how to actually construct this matrix with the maximal determinant.
+It is straight forward to implement the steps outlined in section \ref{sec.procedure}.
+Even for models with a large number of stochastic variables the procedure is both robust and fast.
+
+\vspace{2cm}
+
+\section*{Disclaimer}
+The views and opinions expressed here are those of the authors and do not necessarily represent the views and opinions of their employers.
 
@@
 	\bibitem[Smith, 2008]{smith} Smith, R. L. (2008). The positive definite completion problem revisited. \textit{Linear Algebra and its Applications} 429:1442 -- 1452.	
+    \bibitem[Georgescu, 2018]{georgescu} Georgescu, D.
+I., Higham, N.
+J.
+\& Peters, G.
+W.
+(2018).
+Explicit solutions to correlation matrix completion problems, with an application to risk management and insurance.
+\textit{Roy Soc Open Sci} 5: 172348.
 	\bibitem[Rao, 2002]{rao} C. R. Rao (2002). \textit{Linear Statistical Inference and its Applications, 2nd edition}. John Wiley \& Sons, Inc.
@@
 	\bibitem[Horn, 2005]{horn} Horn, R. A. (2005). Basic  Properties  of  the  Schur  Complement, 17--46. In Fuzhen Zhang (Ed.) \textit{The Schur complement and its application}. Springer.
-	\bibitem[Dreyer et~al., 2021] {ksd} our other paper ...
+	% \bibitem[Dreyer et~al., 2021] {ksd} our other paper ...
 	\bibitem[Golumbic, 2004]{golumbic} Golumbic, M. C. (2004). \textit{Algorithmic Graph Theory and Perfect Graphs}. Elsevier. 

### 2021-11-22 paper/paper_risk.tex
Last small changes

@@
 	We describe a way to complete a correlation matrix that is not fully specified.
-Such matrices arise in financial applications when the number of stochastic variables becomes large or when several smaller models are combined in a larger model.
+Such matrices often arise in financial applications when the number of stochastic variables becomes large or when several smaller models are combined in a larger model.
 We argue that the proper completion to consider is the matrix that maximizes the entropy of the distribution described by the matrix.
@@
 If this graph is chordal our construction will result in a proper correlation matrix.
-We demonstrate the construction using a cross-currency model with six stochastic variables.
+We give a detailed description of the construction for a cross-currency model with six stochastic variables and describe extensions to larger models involving more currencies.
 \end{abstract}
@@
 While it would be desirable to calibrate all these coefficients this is in practice not possible because such a calibration procedure would be too slow, too unstable, or would require calibration instruments for which quotes can not be obtained.
-The issue of stability is especially important in a risk neutral setup.
-Daily calibrations might display undesired fluctuations if the calibration is not stable.
+The issue of stability is especially important in a risk neutral setup where daily calibrations might display undesired fluctuations if the calibration is not stable.
 Hybrid models are another source of undetermined correlation coefficients.
@@
 The most important of which is that one might end up with a matrix that is not positive semidefinite and thus not a valid correlation matrix.
-To obtain a valid matrix further steps are required.
+To obtain a valid matrix, further steps are required.
 One possibility is to use the fact that the set of positive definite correlation matrices is an open set that contains the unit matrix.
@@
 Since we are looking at a correlation matrix, the vertices correspond to the stochastic variables of the model.
-Two vertices are connected by an edge in the graph if the corresponding correlation coefficient in the matrix is know.
-It is here that we encounter the one condition that needs to be satisfied for the construction to work.
-The graph that we have just constructed needs to be chordal.
+Two vertices are connected by an edge in the graph if the corresponding correlation coefficient in the matrix is known.
+It is here that we encounter the one condition that needs to be satisfied for the construction to work: The graph that we have just constructed needs to be chordal.
 A graph is chordal if every loop in the graph that has more than three edges contains a chord, i.e.
@@
 The two interest rate models and the model for the exchange rate are assumed to be stochastic volatility models so that we end up with a total of six stochastic variables.
-We will use this model to describe the all the required steps.
+We will use this model to describe all the required steps.
 These steps can easily be generalized to more complicated models.
@@
 \end{equation}
-be a $k$ dimensional random variable with a density function given by a $k$ dimensional normal distribution
+be a $k$-dimensional random variable with a density function given by a $k$-dimensional normal distribution
 \begin{align}
@@
 
-This choice of $W$ can be characterized in two more ways (see \cite{smith} and \cite{ksd}).
-Banachiewicz showed that the inverse of a matrix can be expressed using the Schur complement.
+This choice of $W$ can be characterized in two more ways (see \cite{smith}).
+Banachiewicz showed that the inverse of a matrix can be expressed using the Schur complement (see \cite{horn}).
 Since the Schur complement $H/C$ is block diagonal the inverse of $H$ is given by
@@
 
-Our model will consist of two interest rate models; one for the domestic currency $E$ and one the foreign currency $A$.
+Our model will consist of two interest rate models; one for the domestic currency $E$ and one for the foreign currency $A$.
 These two currencies are linked by an exchange rate $X$.
@@
 In this paper we have supplied both.
-We have argued that the correct way to complete the matrix is the one that maximizes the entropy of the distribution described by the matrix.
+We have argued that the correct way to complete the matrix is to choose the matrix that maximizes the entropy of the distribution described by the matrix.
 This is the matrix with the maximal determinant.
-We have then shown how to actually construct this matrix with the maximal determinant.
-It is straight forward to implement the steps outlined in section \ref{sec.procedure}.
+Then we have shown how to actually construct this matrix.
+The procedure that we have described in section \ref{sec.procedure} is straight forward to implement and is guaranteed to produce a valid correlation matrix provided one starts with a matrix whose associated graph is chordal.
 Even for models with a large number of stochastic variables the procedure is both robust and fast.
@@
 	\bibitem[Horn, 2005]{horn} Horn, R. A. (2005). Basic  Properties  of  the  Schur  Complement, 17--46. In Fuzhen Zhang (Ed.) \textit{The Schur complement and its application}. Springer.
-	% \bibitem[Dreyer et~al., 2021] {ksd} our other paper ...
 	\bibitem[Golumbic, 2004]{golumbic} Golumbic, M. C. (2004). \textit{Algorithmic Graph Theory and Perfect Graphs}. Elsevier. 

### 2021-11-24 paper/paper_risk.tex
Small changes. Submitted to Risk.

@@
 Modern financial models are driven by a large number of stochastic variables.
-These stochastic variables are usually correlated which creates the need to determine the coefficients of the underlying correlation matrix.
+These normally distributed stochastic variables are usually correlated which creates the need to determine the coefficients of the underlying correlation matrix.
 Since the number of correlation coefficients increases quadratically with the number of stochastic variables, this number can become very large.

### 2021-11-29 paper/paper_math.tex
Intro and abstract for math paper.

@@
 \begin{document}
-\title{Matrix completion and semidefinite matrics}
+\title{Matrix completion and semidefinite matrices}
 
-\author{Olaf Dreyer\thanks{OD Consulting}
-\and Horst K\"ohler \thanks{Commerzbank AG}
-\and Thomas Streuer \thanks{RIVACON GmbH} }
+\author{Olaf Dreyer\thanks{OD Consulting, Frankfurt am Main, email: olaf.dreyer@gmail.com}}
 
@@
 
-\begin{abstract} Here comes the abstract.
+\begin{abstract} Positive semidefinite Hermitian matrices that are not fully specified but whose underlying graph is chordal can be completed to a positive semidefinite Hermitian matrix.
+For positive definite partial Hermitian matrices this completion can be characterized as the unique completion that maximizes the determinant of the matrix.
+It is also the unique matrix whose inverse has zeroes in the places that were undetermined in the original matrix.
+This paper extends these characterizations to the case of semidefinite matrices.
+Because the determinant vanishes for singular matrices and the inverse does not exist we introduce a generalized determinant and use generalized inverses to formulate equivalent characterizations in the semidefinite case.
+For a class of matrices that are singular but of maximal rank unique characterizations can be given just as in the positive definite case.
 \end{abstract}
 
-\emph{AMS classification:} 15A15; 15A18; 15A83; 15B48
+\emph{AMS classification:} 15A10; 15A15; 15A18; 15A83; 15B48
  \\
 
-\emph{Keywords:} positive definite; positive semidefinite; matrix completions;
+\emph{Keywords:} positive definite; positive semidefinite; matrix completions
 
 \section{Introduction}
-And some introduction.
+The question of which partial Hermitian matrices $H$ can be completed to give a fully specified positive definite Hermitian matrix was solved in \cite{grone}.
+The solution provided in \cite{grone} gave one element of the completion at a time.
+While giving the correct solution the practical implementation of this procedure was rather tedious.
+In \cite{smith} a faster procedure was proposed that delivered the same matrix in far fewer steps by focusing on whole blocks at once.
 
-The Schur complement needs to be introduced here.
+	The starting point of both procedures is the graph $\gamma$ that is constructed from the incomplete matrix.
+The vertices of $\gamma$ are given by the rows $i$, $i=1,\ldots,n$, (or columns) of the matrix $H$.
+Two vertices $i$ and $j$ of $\gamma$ are connected by an edge $e$ if the element
+	\begin{equation}
+		a_{ij}
+	\end{equation}
+	of the matrix is given.
+There is now one condition that the graph $\gamma$ has to satisfy for the procedure to work.
+The graph has to be \emph{chordal}.
+A graph is chordal if every loop of four or more elements has a chord, i.e.
+an edge that connects two nonconsecutive elements of the loop (for an introduction to graph theoretic notions see \cite{golumbic}).
+Assuming that the graph is chordal we now proceed by constructing another graph $\Gamma$.
+The vertices of $\Gamma$ are given by the cliques of $\gamma$ (a clique is a maximal set of vertices that are all connected to each other; again see \cite{golumbic}) and two vertices of $\Gamma$ are connected if the two corresponding cliques have a nonempty intersection (see figure \ref{fig.allgraphs}).
+
+\begin{figure}[hbt]
+  \begin{center}
+  	\includegraphics[width=10cm]{img/allGraphs.png}
+  \end{center}
+  \caption{a) Given an incomplete matrix $H$ we can construct a graph $\gamma$ that has an edge for all the specified elements of the matrix.
+The vertices of the graph are the rows (or columns) of the matrix.
+For the completion procedure to work the graph $\gamma$ needs to be chordal.
+The graph shown here is chordal because the central loop only has three elements.
+b) The dashed lines show the cliques of $\gamma$.
+c) The cliques are the vertices of the new graph $\Gamma$.
+Edges in $\Gamma$ connect cliques with a nonempty intersection.}\label{fig.allgraphs}
+\end{figure}
+
+The contribution of \cite{smith} is to show that the completion procedure can be reduced to edges of the graph $\Gamma$.
+Any edge in $\Gamma$ is given by two cliques $\alpha$ and $\beta$ that have a nonzero intersection $\alpha\cap\beta\ne\emptyset$.
+Let
+\begin{equation}
+	H[\alpha] = \begin{pmatrix}
+		A & B \\
+		B^\star & C 
+	\end{pmatrix}
+\end{equation}
+be the submatrix of $H$ with rows and column indices in $\alpha$ and let 
+\begin{equation}
+	H[\beta] = \begin{pmatrix}
+		C & D \\
+		D^\star & E 
+	\end{pmatrix}
+\end{equation}
+be the submatrix of $H$ with rows and column indices in $\beta$. We then have
+\begin{equation}
+	H[\alpha\cap\beta] = C.
+\end{equation}
+We can combine these two matrices in one large matrix with row and column indices in $\alpha\cup\beta$,
+\begin{equation}
+	H[\alpha\cup\beta] = \begin{pmatrix}
+		A & B & X \\
+		B^\star & C & D \\
+		X^\star & D^\star & E
+	\end{pmatrix},
+\end{equation}
+where the matrix $X$ remains to be specified.
+It was now shown in \cite[Theorem 3.2.]{smith} that a partial positive semidefinite matrix $H[\alpha\cup\beta]$ can be completed to a positive semidefinite matrix by choosing
+\begin{equation}\label{eqn.defx}
+	X = B C^+ D,
+\end{equation}
+where $C^+$ denotes the Moore-Penrose inverse of $C$ (see \cite{roger} and \cite{matrixanalysis}).
+Note that the matrix does not need to be positive definite for the construction to work.
+It it sufficient for the matrix to be positive \emph{semi}definite.
+
+If the matrix is positive definite the choice of $X$ in equation (\ref{eqn.defx}) can be uniquely characterized in two ways.
+First it is the choice that maximizes the determinant of the completed matrix.
+It is also the choice for which the inverse of $H[\alpha\cup\beta]$ has zeroes in those places where $X$ used to be.
+
+We see that the treatment of positive definite and positive semidefinite matrices differs.
+For positive definite matrices the completion in equation (\ref{eqn.defx}) has properties that uniquely determine it.
+For the semidefinite case we are just left with the existence result.
+Since the determinant of a singular matrix vanishes and since the inverse of a singular matrix does not exist it is not clear if we can hope to improve this situation.
+
+In this paper show we now make two contributions.
+The first contribution is a new prove that the $X$ from equation (\ref{eqn.defx}) gives a positive semidefinite completion of $H[\alpha\cup\beta]$.
+We then show that we can extend the uniqueness results to the semidefinite case.
+To do this we need to introduce a generalized determinant that gives the determinant of the nonsingular part of the matrix.
+This determinant does not have many of the nice properties that the usual determinant has.
+If we restrict our attention to a special class of matrices though we can prove three results that hold for positive definite matrices to the case of positive semidefinite matrices: Fischer's inequality, the Schur determinant lemma, and Banachiewicz's form of the inverse.
+These three results allow us to extend the uniqueness results to semidefinite matrices.
+We start by introducing the class of matrices for which our results will be valid.
 
@@
 \begin{definition}
-	Let $H$ be a positive semidefinite Hermitian matrix that is partitioned as in equation (\ref{eqn.defh}). We say that $H$ is of \emph{maximal rank} if and only if $N(H) = N(A)\oplus N(C)$. 
+	Let $H$ be a positive semidefinite Hermitian matrix that is partitioned as in equation (\ref{eqn.defh}).
+We say that $H$ is of \emph{maximal rank} if and only if $N(H) = N(A)\oplus N(C)$\footnote{We should be precise and say that $H$ is positive semidefinite with respect to the partition in equation (\ref{eqn.defh}).
+We will refrain from doing this and rely on the reader to infer the partition from the context.}.
 \end{definition}
@@
 \end{equation} 
-We will use this property to extend results that are valid for positive definite matrices to partitioned matrices of maximal rank.
-Before we can state these results we need to introduce one more notion.
+in section \ref{sec.ext} we will use this property to extend results that are valid for positive definite matrices to partitioned matrices of maximal rank.
+We need one more notion before we can formulate these results.
 
 \section{The generalized determinant}\label{sec.gendet}
-A Hermitian matrix $H$ defines a non-singular map from its range $R(H)$ to its range $R(H)$.
-If $H$ is singular its determinant vanishes.
-The determinant thus contains no information about $H$ restricted to $R(H)$.
+A Hermitian matrix $H$ defines a nonsingular map from its range $R(H)$ to its range $R(H)$.
+If the nullspace $N(H)$ is nonzero, i.e.
+if $H$ is singular, the determinant of $H$ vanishes.
+The determinant thus contains no information about the nonsingular map that is $H$ restricted to $R(H)$.
 To recover this information we introduce a generalized determinant $\detp$:
@@
 
-We note that the generalized determinant does not have many of the properties the usual determinant has.
+We note that the generalized determinant lacks many of the properties the usual determinant has.
 In particular, it is not a continuous function of $H$, and the generalized determinant of the product of two matrices is not the product of the two generalized determinants.
@@
 \section{The extensions}\label{sec.ext}
-With the preparations in section \ref{sec.partition}, and the definition of the generalized determinant in the last section, we now want to extent three results: Fischer's inequality, the determinant equality for Schur's complement, and Banachiewicz's form of the inverse of a matrix.
+With the preparations in section \ref{sec.partition}, and the definition of the generalized determinant in the last section, we now want to extent three results to singular matrices: Fischer's inequality, the determinant equality for the Schur complement, and Banachiewicz's form of the inverse of a matrix.
 
@@
 If $H$ is singular, the determinant of $H$ vanishes and the inequality is no longer much of a constraint.
-We can also no longer infer that $B=0$ in the case of equality.
+We can also no longer infer that $B$ vanishes in the case of equality.
 
@@
 \end{equation}
-We now look at the restriction of $H$ and all its sub-matrices to this space and use the usual Fischer inequality there.
+We now look at the restriction of $H$ and all its submatrices to this space and use the Fischer's inequality there.
+Note that because $H$ is positive semidefinite it has the column (and row) inclusion property and we have
+\begin{equation}
+	N(A) \subset N(B^\star) 
+\end{equation}
+and
+\begin{equation}
+	R(B^\star) \subset R(C). 
+\end{equation}
+$B^\star$ thus vanishes on the complement of $R(A)$ and maps into $R(C)$.
+The restriction of $B^\star$ to $R(A)$ thus keeps the nontrivial part of $B^\star$.
+The same is true for $B$ and its restriction to $R(C)$.
 As in definition \ref{def.gendet}, we denote the restriction of $H$ to its range by $\bar H$.
@@
 \subsection{Schur complement}
-We now turn our attention to the Schur complement. For a positive definite $H$ Schur showed that
+The Schur complement arises naturally when using Gaussian elimination to solve a linear equation. Let $H$ be as in equation (\ref{eqn.defh}) and let us assume we want to solve
+\begin{align}
+0 & = H \begin{pmatrix}
+		k \\
+		l
+	\end{pmatrix} \\
+	& = \begin{pmatrix}
+		A k + B l \\
+		B^\star k + C l
+ \end{pmatrix}	\label{eqn:lineq}
+\end{align}
+Assuming that $A$ is invertible we can solve for $k$ in the first equation of (\ref{eqn:lineq}) to obtain
+\begin{equation}\label{eqn.gauss1}
+	k = - A^{-1}B l.
+\end{equation}
+The second equation of (\ref{eqn:lineq}) then gives
+\begin{equation}\label{eqn.gauss2}
+	(C - B^\star A^{-1}B) l = 0.
+\end{equation}
+The expression in the parentheses is called the Schur complement $H/A$ of $A$ in $H$:
+\begin{equation}
+	H/A = C - B^\star A^{-1}B
+\end{equation}
+For a positive semidefinite matrix $H$ we may generalize the definition to
+\begin{equation}
+	H/A = C - B^\star A^+B,
+\end{equation}
+where $A^+$ is the Moore-Penrose inverse of $A$ (see \cite{roger} and \cite{matrixanalysis}).
+Note that the choice of generalized inverse does not matter here since $R(B)\subset R(A)$.
+Emilie Haynsworth introduced the name and showed that the Schur complement possesses many interesting properties \cite{hayns} (see \cite{horn} for more details).
+In \cite{schur} Issai Schur showed that for a positive definite $H$ the determinant of $H$ satisfies
 \begin{equation}
@@
 \end{equation}
-We again focus our attention on $R(A)\oplus R(C)$ where $H$ is positive definite. Using the notation from the previous section we obtain:
+We want to show now that this equality holds also for the generalized determinant.
+Again, we focus our attention on $R(A)\oplus R(C)$ where $H$ is positive definite.
+Using the notation from the previous section we obtain:
 \begin{align}
@@
 \end{align}
-We already know that $N(C) \subset N(H/A)$. To show equality let $l\in N(H/A)$ and set
+We already know that $N(C) \subset N(H/A)$.
+To show equality let $l\in N(H/A)$.
+This $l$ thus satisfies equation (\ref{eqn.gauss2}) that was obtained through Gaussian elimination.
+We define $k$ according to equation (\ref{eqn.gauss1}):
 \begin{equation}
@@
 \subsection{The inverse}
-The result concerns the inverse of $H$. When $H$ is positive definite its inverse can be written as
+The last result concerns the inverse of $H$. When $H$ is positive definite its inverse can be written as
 \begin{equation}\label{eqn.bana}
@@
 \end{equation} 
-This is a remarkable formula because to find the of $H$ we just need to invert $A$ and $H/A$.
+This is a remarkable formula because to find the inverse of $H$ we just need to invert $A$ and $H/A$.
 This formula was first established by Banachiewicz \cite{bana} (see also \cite[p.112]{frazer}).
@@
 \end{equation}
-To denote sub-matrices of $H$ we use the notation from \cite{matrixanalysis}. For a set $\alpha\subset \{1, \ldots, n\}$ we let
+To denote submatrices of $H$ we use the notation from \cite{matrixanalysis}. For a set $\alpha\subset \{1, \ldots, n\}$ we let
 \begin{equation}
@@
 \end{equation}
-be the sub-matrix of $H$ with row and column indices in $\alpha$. For $\alpha,\beta\subset \{1, \ldots, n\}$ we let
+be the submatrix of $H$ with row and column indices in $\alpha$. For $\alpha,\beta\subset \{1, \ldots, n\}$ we let
 \begin{equation}
@@
 \end{equation}
-be the sub-matrix with row indices in $\alpha$ and column indices in $\beta$. Now let $\alpha$ and $\beta$ be such that
+be the submatrix with row indices in $\alpha$ and column indices in $\beta$. Now let $\alpha$ and $\beta$ be such that
 \begin{equation}
@@
 The first result characterizes $X$ as the unique extension of $H$ that maximizes the determinant of $H$.
-If $H$ is non-singular we can just talk about the regular determinant of $H$.
+If $H$ is nonsingular we can just talk about the regular determinant of $H$.
 If $H$ is singular we have to use the generalized determinant that we introduced in section \ref{sec.gendet}.
@@
 
-For a non-singular matrix $H$ we can use the determinant to find the inverse of $H$ (see \cite{matrixanalysis}):
+For a nonsingular matrix $H$ we can use the determinant to find the inverse of $H$ (see \cite{matrixanalysis}):
 \begin{equation}\label{eqn.inverseformula}
@@
 \end{equation}
-Because we are expressing $X$ in terms of the given sub-matrices $A, \ldots, E$ of $H$, it is clear that the derivative in equation (\ref{eqn.inverseformula}) vanishes for indices $i$ and $j$ that denote elements of $X$ itself.
-$H^{-1}$ has zeroes in those places where the matrix $X$ sits in $H$.
+
+Since $X$ was chosen such that the determinant is maximal the derivative in equation (\ref{eqn.inverseformula}) vanishes for indices $i$ and $j$ that denote elements of $X$ itself.
+It follows that $H^{-1}$ has zeroes in those places where the matrix $X$ sits in $H$.
 It turns out that this characteristic uniquely determines $X$, even if $H$ is only positive semidefinite and we have to talk about the Moore-Penrose inverse of $H$ instead.
@@
 	Because $H$ is of maximal rank we can use proposition \ref{prop.bana} to express the Moore-Penrose inverse of $H$ in terms of $C^+$ and $(H/C)^+$.
-If $H^+$ has zeroes where $X$ is in $H$ then we must have
+If $H^+$ is to have zeroes where $X$ is in $H$ then we must have
 	\begin{equation}
@@
 \end{equation}
+
 \section{Conclusion}
 
+\section*{Acknowledgement}
+I would like to thank Patrick B\"uchel for his support during the creation of this work as well as Horst K\"ohler and Thomas Streuer for their initial push to look into maximal determinant completions of matrices.
+
 \begin{thebibliography}{MM}
+	\bibitem{grone} R. Grone, C. R. Johnson, E. Sa, H. Wolkowicz, Positive definite completions of partial Hermitian matrices, Linear Algebra Appl. 58 (1984) 109-124.
+	\bibitem{smith} Smith, R. (2008). The positive definite completion problem revisited Linear Algebra and its Applications  429(7), 1442-1452. https://dx.doi.org/10.1016/j.laa.2008.04.020
+	\bibitem{golumbic} M.C. Golumbic, W. Rheinboldt, Algorithmic Graph Theory and Perfect Graphs, 1st edition, Elsevier Inc, Academic Press, 1980.
+	\bibitem{roger} Penrose, R. (1955). A generalized inverse for matrices Mathematical Proceedings of the Cambridge Philosophical Society  51(3), 406-413. https://dx.doi.org/10.1017/s0305004100030401
 	\bibitem{matrixanalysis} Roger A. Horn, Charles R. Johnson, \emph{Matrix Analysis}, 2nd Edition, Cambridge University Press, 2013.
+	\bibitem{hayns} Emilie Virginia Haynsworth, Determination of the inertia of a partitioned Hermitian matrix,  Linear  Algebra and its Applications \textbf{1}, 1968, 73--81.
+	\bibitem{horn} Roger A. Horn, Basic  Properties  of  the  Schur  Complement. In Fuzhen Zhang (Ed.) The Schur complement and its application (17--46). Springer, 2005.
 	\bibitem{schur} J. Schur, Über Potenzreihen, die im Innern des Einheitskreises beschränkt sind, Journal für die reine und angewandte Mathematik \textbf{147}, 205 -- 232, 1917.
@@
 https://dx.doi.org/10.1137/0117041
-	\bibitem{smith} Smith, R. (2008). The positive definite completion problem revisited Linear Algebra and its Applications  429(7), 1442-1452. https://dx.doi.org/10.1016/j.laa.2008.04.020	
-	% older stuff
-	\bibitem{hayns} Emilie Virginia Haynsworth, Determination of the inertia of a partitioned Hermitian matrix,  Linear  Algebra and its Applications \textbf{1}, 1968, 73--81.
-	\bibitem{horn} Roger A. Horn, Basic  Properties  of  the  Schur  Complement. In Fuzhen Zhang (Ed.) The Schur complement and its application (17--46). Springer, 2005.
-	\bibitem{rao} C. Radharkishna Rao, Linear Statistical Inference and its Applications, 2nd edition, John Wiley \& Sons, Inc, 2002.
-	\bibitem{cottle} Richard W. Cottle, Manifestations of the Schur complement, Linear Algebra and its Applications \textbf{8}, 1974, 189--211. 
-	\bibitem{roger} Penrose, R. (1955). A generalized inverse for matrices Mathematical Proceedings of the Cambridge Philosophical Society  51(3), 406-413. https://dx.doi.org/10.1017/s0305004100030401
 \end{thebibliography}

### 2021-12-01 paper/paper_math.tex
All the pieces in place.

@@
 
-\begin{abstract} Positive semidefinite Hermitian matrices that are not fully specified but whose underlying graph is chordal can be completed to a positive semidefinite Hermitian matrix.
-For positive definite partial Hermitian matrices this completion can be characterized as the unique completion that maximizes the determinant of the matrix.
-It is also the unique matrix whose inverse has zeroes in the places that were undetermined in the original matrix.
-This paper extends these characterizations to the case of semidefinite matrices.
-Because the determinant vanishes for singular matrices and the inverse does not exist we introduce a generalized determinant and use generalized inverses to formulate equivalent characterizations in the semidefinite case.
+\begin{abstract} For positive semidefinite Hermitian matrices that are not fully specified can be completed to a positive definite Hermitian matrix provided the underlying graph of the matrix is chordal.
+For positive definite partial Hermitian matrices certain unique completions can be singled out.
+The completion that maximizes the determinant of the matrix is unique.
+This is also the unique matrix whose inverse has zeroes in those places that were undetermined in the original matrix.
+This paper extends these uniqueness results to the case of semidefinite matrices.
+Because the determinant vanishes for singular matrices and because the inverse does not exist we introduce a generalized determinant and use generalized inverses to formulate equivalent characterizations in the semidefinite case.
 For a class of matrices that are singular but of maximal rank unique characterizations can be given just as in the positive definite case.
@@
 The question of which partial Hermitian matrices $H$ can be completed to give a fully specified positive definite Hermitian matrix was solved in \cite{grone}.
-The solution provided in \cite{grone} gave one element of the completion at a time.
+The solution provided in \cite{grone} was constructive but only gave one element of the completion with each step.
 While giving the correct solution the practical implementation of this procedure was rather tedious.
-In \cite{smith} a faster procedure was proposed that delivered the same matrix in far fewer steps by focusing on whole blocks at once.
+In \cite{smith} a faster procedure was proposed that delivered the same matrix in far fewer steps by focusing on whole blocks of the matrix at once.
 
-	The starting point of both procedures is the graph $\gamma$ that is constructed from the incomplete matrix.
-The vertices of $\gamma$ are given by the rows $i$, $i=1,\ldots,n$, (or columns) of the matrix $H$.
+The starting point of both procedures is the graph $\gamma$ that is constructed from the partial matrix.
+The vertices of $\gamma$ are given by the rows (or columns) $i$, $i=1,\ldots,n$, of the matrix $H$.
 Two vertices $i$ and $j$ of $\gamma$ are connected by an edge $e$ if the element
 	\begin{equation}
-		a_{ij}
+		h_{ij}
 	\end{equation}
 of the matrix is given.
-There is now one condition that the graph $\gamma$ has to satisfy for the procedure to work.
-The graph has to be \emph{chordal}.
+There is now one condition that the graph $\gamma$ has to satisfy for the procedure to work: The graph has to be \emph{chordal}.
 A graph is chordal if every loop of four or more elements has a chord, i.e.
 an edge that connects two nonconsecutive elements of the loop (for an introduction to graph theoretic notions see \cite{golumbic}).
-Assuming that the graph is chordal we now proceed by constructing another graph $\Gamma$.
+Given the graph $\gamma$ we can now proceed by constructing another graph $\Gamma$.
 The vertices of $\Gamma$ are given by the cliques of $\gamma$ (a clique is a maximal set of vertices that are all connected to each other; again see \cite{golumbic}) and two vertices of $\Gamma$ are connected if the two corresponding cliques have a nonempty intersection (see figure \ref{fig.allgraphs}).
@@
   \begin{center}
-  	\includegraphics[width=10cm]{img/allGraphs.png}
+  	\includegraphics[width=12cm]{img/allGraphs.png}
   \end{center}
-  \caption{a) Given an incomplete matrix $H$ we can construct a graph $\gamma$ that has an edge for all the specified elements of the matrix.
+  \caption{a) Given a partial matrix $H$ we can construct a graph $\gamma$ that has an edge for all the specified elements of the matrix.
 The vertices of the graph are the rows (or columns) of the matrix.
 For the completion procedure to work the graph $\gamma$ needs to be chordal.
-The graph shown here is chordal because the central loop only has three elements.
+The graph shown here is chordal because the central loop only has three elements and thus does not require a chord.
 b) The dashed lines show the cliques of $\gamma$.
@@
 The contribution of \cite{smith} is to show that the completion procedure can be reduced to edges of the graph $\Gamma$.
-Any edge in $\Gamma$ is given by two cliques $\alpha$ and $\beta$ that have a nonzero intersection $\alpha\cap\beta\ne\emptyset$.
+Any edge in $\Gamma$ is given by two cliques $\alpha$ and $\beta$ that have a nonzero intersection $\alpha\cap\beta$.
 Let
@@
 \end{equation}
-where $C^+$ denotes the Moore-Penrose inverse of $C$ (see \cite{roger} and \cite{matrixanalysis}).
+where $C^+$ denotes the Moore-Penrose inverse of $C$ (see \cite{roger} and \cite{matrixanalysis} for details on the Moore-Penrose inverse).
 Note that the matrix does not need to be positive definite for the construction to work.
@@
 If the matrix is positive definite the choice of $X$ in equation (\ref{eqn.defx}) can be uniquely characterized in two ways.
-First it is the choice that maximizes the determinant of the completed matrix.
-It is also the choice for which the inverse of $H[\alpha\cup\beta]$ has zeroes in those places where $X$ used to be.
+First, it is the choice that maximizes the determinant of the completed matrix.
+It is also the choice for which the inverse of $H[\alpha\cup\beta]$ has zeroes in those places where $X$ sits in $H[\alpha\cup\beta]$.
 
@@
 
-In this paper show we now make two contributions.
-The first contribution is a new prove that the $X$ from equation (\ref{eqn.defx}) gives a positive semidefinite completion of $H[\alpha\cup\beta]$.
-We then show that we can extend the uniqueness results to the semidefinite case.
+In this paper we now make two contributions.
+The first contribution is a new proof that the $X$ from equation (\ref{eqn.defx}) gives a positive semidefinite completion of $H[\alpha\cup\beta]$.
+We then extend the uniqueness results to the semidefinite case.
 To do this we need to introduce a generalized determinant that gives the determinant of the nonsingular part of the matrix.
-This determinant does not have many of the nice properties that the usual determinant has.
-If we restrict our attention to a special class of matrices though we can prove three results that hold for positive definite matrices to the case of positive semidefinite matrices: Fischer's inequality, the Schur determinant lemma, and Banachiewicz's form of the inverse.
-These three results allow us to extend the uniqueness results to semidefinite matrices.
-We start by introducing the class of matrices for which our results will be valid.
+This determinant lacks many of the nice properties that the usual determinant has.
+If we restrict our attention to a special class of matrices though we can prove three results that hold for positive definite matrices: Fischer's inequality, the Schur determinant lemma, and Banachiewicz's form of the inverse.
+These three results will then be used to uniquely characterize the completion in equation (\ref{eqn.defx}) for semidefinite matrices.
+
+We start by introducing the class of matrices for which our results are valid.
 
@@
 \end{equation}
-In a similar fashion we can establish that the null space of $C$ is contained in the null space of $B$:
+In a similar fashion we can establish that the nullspace of $C$ is contained in the null space of $B$:
 \begin{equation}\label{eqn.nullc}
@@
 \end{equation}
-The relations for the null spaces are equivalent to these relations for the ranges of $A$ and $C$:
+The relations for the nullspaces are equivalent to these relations for the ranges of $A$ and $C$:
 \begin{align}
@@
 We have equality if and only if $N(H) = N(A)\oplus N(C)$.
-In the following, matrices $H$ for which this equality holds, will be of particular interest to us which is why we make the following definition:
+In the following, matrices $H$ for which this equality holds will be of particular interest to us which is why we make the following definition:
 
@@
 	Let $H$ be a positive semidefinite Hermitian matrix that is partitioned as in equation (\ref{eqn.defh}).
-We say that $H$ is of \emph{maximal rank} if and only if $N(H) = N(A)\oplus N(C)$\footnote{We should be precise and say that $H$ is positive semidefinite with respect to the partition in equation (\ref{eqn.defh}).
-We will refrain from doing this and rely on the reader to infer the partition from the context.}.
+We say that $H$ is of \emph{maximal rank} if and only if $N(H) = N(A)\oplus N(C)$\footnote{We should be precise and say that $H$ is of maximal rank with respect to the partition in equation (\ref{eqn.defh}).
+For the sake of readability we will refrain from doing so and rely on the reader to infer the partition from the context.}.
 \end{definition}
@@
 \end{equation} 
-in section \ref{sec.ext} we will use this property to extend results that are valid for positive definite matrices to partitioned matrices of maximal rank.
+In section \ref{sec.ext} we will use this property to extend results that are valid for positive definite matrices to partitioned matrices of maximal rank.
 We need one more notion before we can formulate these results.
@@
 
-\begin{lemma}
+\begin{lemma}\label{lem.gentdetprop}
 	Let $H$ be Hermitian of rank $r\le n$. Let $\lambda_i, i=1, \ldots, n$, be the eigenvalues of $H$. Let us assume that they are ordered in such a way that $\lambda_i = 0$, for $i>r$. We then have:
@@
 		\item We have
-		\begin{equation}
+		\begin{equation}\label{eqn.limit}
 			\detp H = \lim_{\epsilon \rightarrow 0} \frac{\det( H + \epsilon I )}{\epsilon^{n-r}}
@@
 
+To show the inequality we could have also used equation (\ref{eqn.limit}) of lemma \ref{lem.gentdetprop}.
+
 \subsection{Schur complement}
@@
 		B^\star k + C l
- \end{pmatrix}	\label{eqn:lineq}
+ \end{pmatrix}.	\label{eqn:lineq}
 \end{align}
@@
 \end{equation}
-where $A^+$ is the Moore-Penrose inverse of $A$ (see \cite{roger} and \cite{matrixanalysis}).
+where $A^+$ is the Moore-Penrose inverse of $A$ (again, see \cite{roger} and \cite{matrixanalysis}).
 Note that the choice of generalized inverse does not matter here since $R(B)\subset R(A)$.
-Emilie Haynsworth introduced the name and showed that the Schur complement possesses many interesting properties \cite{hayns} (see \cite{horn} for more details).
+Emilie Haynsworth introduced the name and showed that the Schur complement possesses many interesting properties \cite{hayns} (see \cite{horn} for a detailed exposition).
 In \cite{schur} Issai Schur showed that for a positive definite $H$ the determinant of $H$ satisfies
@@
 \end{equation}
-We want to show now that this equality holds also for the generalized determinant.
+We now want to show that this equality also holds for the generalized determinant.
 Again, we focus our attention on $R(A)\oplus R(C)$ where $H$ is positive definite.
@@
 \end{align}
-We need to convince ourselves that the last determinant is equal to $\detp H/A$. To check this we need to show that 
+We need to convince ourselves that the last determinant is equal to $\detp H/A$. To check this, we need to show that 
 \begin{align}
@@
 $\bar H$ is positive definite and we can express its inverse in the form of equation (\ref{eqn.bana}).
-In the last section we established that the range and null spaces of $C$ and $H/A$ coincide so that by replacing the inverses with Moore-Penrose inverses we obtain the unique Moore-Penrose inverse of $H$.
+In the last section we established that the ranges and nullspaces of $C$ and $H/A$ coincide so that by replacing the inverses in equation (\ref{eqn.bana}) with Moore-Penrose inverses we obtain the unique Moore-Penrose inverse of $H$.
 
@@
 
-This result can also be found in \cite{ouellette}.
+This result can also be deduced from \cite[Theorem 4.6]{ouellette}.
 
@@
 \end{equation}
-To denote submatrices of $H$ we use the notation from \cite{matrixanalysis}. For a set $\alpha\subset \{1, \ldots, n\}$ we let
+As in the introduction we use the notation from \cite{matrixanalysis} to denote submatrices of $H$. For $\alpha\subset \{1, \ldots, n\}$ let
 \begin{equation}
@@
 \begin{equation}
-	\detp H/C \le \detp H[\alpha]/C + \detp H[\beta]/C,
+	\detp H/C \le \detp H[\alpha]/C \detp H[\beta]/C,
 \end{equation}
@@
 \end{equation}
-It follows that this $X$ is the unique choice that maximizes the determinant.
+It follows that this $X$ is the unique choice that maximizes the determinant of $H$.
 \end{proof}
@@
 It follows that $H^{-1}$ has zeroes in those places where the matrix $X$ sits in $H$.
-It turns out that this characteristic uniquely determines $X$, even if $H$ is only positive semidefinite and we have to talk about the Moore-Penrose inverse of $H$ instead.
+It turns out that this uniquely determines $X$ even if $H$ is only positive semidefinite and we have to talk about the Moore-Penrose inverse of $H$ instead.
 
@@
 \section{Conclusion}
+The problem of how to complete partial Hermitian matrices arises frequently in practical applications (see \cite{otherpaper} for an example from finance).
+This problem was solved in \cite{grone} for partial matrices whose corresponding graph is chordal.
+The procedure provided in \cite{grone} was improved in \cite{smith} by giving a way to calculate whole blocks of the completion at once.
+For positive definite matrices the resulting completion is singled out by two uniqueness results.
+It is the unique matrix that maximizes the determinant, and it is the unique matrix whose inverse has zeroes in those places that were unspecified in the original matrix.
+In this paper we have extended these uniqueness results to include semidefinite matrices.
+To make this extension possible we needed to introduce a generalized determinant that gives the determinant of the nontrivial part of a Hermitian matrix.
+We also needed to focus on matrices whose rank is determined solely by the rank of its diagonal matrices.
+For these matrices the same uniqueness results hold that hold for positive definite matrices.
 
 \section*{Acknowledgement}
-I would like to thank Patrick B\"uchel for his support during the creation of this work as well as Horst K\"ohler and Thomas Streuer for their initial push to look into maximal determinant completions of matrices.
+I would like to thank Patrick B\"uchel for his support during the creation of this work as well as Horst K\"ohler and Thomas Streuer for their initial push to look into maximal determinant completions of matrices, and for their support during the creation of this work.
 
@@
 https://dx.doi.org/10.1137/0117041
+	\bibitem{otherpaper} O. Dreyer, H. Köhler, T. Streuer, Completing correlation matrices, arXiv:2111.12640 
 \end{thebibliography}
 
-\appendix
-
-\section{Maybe ...}
-
 \end{document}

### 2021-12-01 paper/paper_math.tex
Fixed bibliography

@@
 \begin{thebibliography}{MM}
-	\bibitem{grone} R. Grone, C. R. Johnson, E. Sa, H. Wolkowicz, Positive definite completions of partial Hermitian matrices, Linear Algebra Appl. 58 (1984) 109-124.
-	\bibitem{smith} Smith, R. (2008). The positive definite completion problem revisited Linear Algebra and its Applications  429(7), 1442-1452. https://dx.doi.org/10.1016/j.laa.2008.04.020
+	\bibitem{grone} R.
+Grone, C.
+R.
+Johnson, E.
+Sa, H.
+Wolkowicz, Positive definite completions of partial Hermitian matrices, Linear Algebra Appl.
+58 (1984) 109--124.
+https://doi.org/10.1016/0024-3795(84)90207-6.
+	\bibitem{smith} R. L. Smith, The positive definite completion problem revisited, Linear Algebra Appl. 429 (2008) 1442--1452. https://dx.doi.org/10.1016/j.laa.2008.04.020
 	\bibitem{golumbic} M.C. Golumbic, W. Rheinboldt, Algorithmic Graph Theory and Perfect Graphs, 1st edition, Elsevier Inc, Academic Press, 1980.
-	\bibitem{roger} Penrose, R. (1955). A generalized inverse for matrices Mathematical Proceedings of the Cambridge Philosophical Society  51(3), 406-413. https://dx.doi.org/10.1017/s0305004100030401
-	\bibitem{matrixanalysis} Roger A. Horn, Charles R. Johnson, \emph{Matrix Analysis}, 2nd Edition, Cambridge University Press, 2013.
-	\bibitem{hayns} Emilie Virginia Haynsworth, Determination of the inertia of a partitioned Hermitian matrix,  Linear  Algebra and its Applications \textbf{1}, 1968, 73--81.
-	\bibitem{horn} Roger A. Horn, Basic  Properties  of  the  Schur  Complement. In Fuzhen Zhang (Ed.) The Schur complement and its application (17--46). Springer, 2005.
-	\bibitem{schur} J. Schur, Über Potenzreihen, die im Innern des Einheitskreises beschränkt sind, Journal für die reine und angewandte Mathematik \textbf{147}, 205 -- 232, 1917.
-	\bibitem{bana} T. Banachiewicz, Zur Berechnung der Determinanten, wie auch der Inversen, und zur darauf basierten Auflösung der Systeme linearer Gleichungen, Acta Astronom. S\'er. C 3, 41--67, 1937.
-	\bibitem{frazer} Frazer, R., Duncan, W., Collar, A. (1938). Elementary Matrices And Some Applications To Dynamics And Differential Equations, Cambridge University Press, 1938.
-	\bibitem{ouellette} Ouellette, D. (1981). Schur complements and statistics Linear Algebra and its Applications  36, 187-295. https://dx.doi.org/10.1016/0024-3795(81)90232-9
-	\bibitem{albert} Albert, A.
-(1969).
-Conditions for Positive and Nonnegative Definiteness in Terms of Pseudoinverses SIAM Journal on Applied Mathematics  17(2), 434-440.
+	\bibitem{roger} R. Penrose, A generalized inverse for matrices, Math Proc Cambridge 51 (1955) 406--413. https://doi.org/10.1017/s0305004100030401.
+	\bibitem{matrixanalysis} R. A. Horn, C. R. Johnson, Matrix Analysis, 2nd ed., Cambridge University Press, 2012.
+	\bibitem{hayns} E. V. Haynsworth, Determination of the inertia of a partitioned Hermitian matrix,  Linear Algebra Appl. 1 (1968) 73--81.
+	\bibitem{horn} R. A. Horn, Basic  Properties  of  the  Schur  Complement. In Fuzhen Zhang (Ed.), The Schur complement and its application (17--46), Springer, 2005.
+	\bibitem{schur} J. Schur, Über Potenzreihen, die im Innern des Einheitskreises beschränkt sind, Journal für die reine und angewandte Mathematik 147 (1917) 205 -- 232.
+	\bibitem{bana} T. Banachiewicz, Zur Berechnung der Determinanten, wie auch der Inversen, und zur darauf basierten Auflösung der Systeme linearer Gleichungen, Acta Astronom. S\'er. C 3 (1937) 41--67.	
+	\bibitem{frazer} R. Frazer, W. Duncan, A. Collar, Elementary Matrices And Some Applications To Dynamics And Differential Equations, Cambridge University Press, 1938.
+	\bibitem{ouellette} D. Ouellette, Schur complements and statistics Linear Algebra Appl. 36 (1981) 187--295. https://dx.doi.org/10.1016/0024-3795(81)90232-9.
+	\bibitem{albert} A.
+Albert, Conditions for Positive and Nonnegative Definiteness in Terms of Pseudoinverses, SIAM Journal on Applied Mathematics  17 (1969) 434--440.
 https://dx.doi.org/10.1137/0117041

### 2021-12-01 paper/paper_math.tex
Small changes to the final version.

@@
 
-\begin{abstract} For positive semidefinite Hermitian matrices that are not fully specified can be completed to a positive definite Hermitian matrix provided the underlying graph of the matrix is chordal.
-For positive definite partial Hermitian matrices certain unique completions can be singled out.
-The completion that maximizes the determinant of the matrix is unique.
-This is also the unique matrix whose inverse has zeroes in those places that were undetermined in the original matrix.
+\begin{abstract} Positive semidefinite Hermitian matrices that are not fully specified can be completed to a positive semidefinite Hermitian matrix provided their underlying graph is chordal.
+If the matrix is positive definite the completion can be uniquely characterized as the matrix that maximizes the determinant, or as the matrix whose inverse has zeroes in those places that were undetermined in the original matrix.
 This paper extends these uniqueness results to the case of semidefinite matrices.
-Because the determinant vanishes for singular matrices and because the inverse does not exist we introduce a generalized determinant and use generalized inverses to formulate equivalent characterizations in the semidefinite case.
-For a class of matrices that are singular but of maximal rank unique characterizations can be given just as in the positive definite case.
+Because the determinant vanishes for singular matrices, and because the inverse does not exist, we introduce a generalized determinant and use generalized inverses to formulate equivalent characterizations in the semidefinite case.
+For a class of matrices that are singular but of maximal rank unique characterizations can be given, just as in the positive definite case.
 \end{abstract}
@@
 The solution provided in \cite{grone} was constructive but only gave one element of the completion with each step.
-While giving the correct solution the practical implementation of this procedure was rather tedious.
+While giving the correct solution, the practical implementation of this procedure was rather tedious.
 In \cite{smith} a faster procedure was proposed that delivered the same matrix in far fewer steps by focusing on whole blocks of the matrix at once.
@@
 This determinant lacks many of the nice properties that the usual determinant has.
-If we restrict our attention to a special class of matrices though we can prove three results that hold for positive definite matrices: Fischer's inequality, the Schur determinant lemma, and Banachiewicz's form of the inverse.
+If we restrict our attention to a special class of matrices, though, we can prove three results that hold for positive definite matrices: Fischer's inequality, the Schur determinant lemma, and Banachiewicz's form of the inverse.
 These three results will then be used to uniquely characterize the completion in equation (\ref{eqn.defx}) for semidefinite matrices.
@@
 \end{align}
-Because we are looking at the restriction of $H$ to $R(A)\oplus R(C)$ and because it is positive definite on $R(A)\oplus R(C)$, we also get that $B$, when restricted to $R(C)$, vanishes if and only if equality holds above.
+Because we are looking at the restriction of $H$ to $R(A)\oplus R(C)$ and because $H$ is positive definite on $R(A)\oplus R(C)$, we also get that $B$, when restricted to $R(C)$, vanishes if and only if equality holds above.
 Since $B$ vanishes on $N(C)$, this is the case if and only if
@@
 
-To show the inequality we could have also used equation (\ref{eqn.limit}) of lemma \ref{lem.gentdetprop}.
+We note that to show the inequality we could have also used equation (\ref{eqn.limit}) of lemma \ref{lem.gentdetprop}.
 
@@
 \end{equation}
-The expression in the parentheses is called the Schur complement $H/A$ of $A$ in $H$:
+The expression in the parentheses is called the Schur complement of $A$ in $H$ and is denoted by $H/A$:
 \begin{equation}
@@
 	\end{equation}
-	is the unique choice for $X$ for which $H$ is positive semidefinite, of maximal rank and for which the (generalized) determinant is maximal.  
+	is the unique choice for $X$ for which $H$ is positive semidefinite, of maximal rank, and for which the (generalized) determinant is maximal.  
 \end{theorem}
@@
 This problem was solved in \cite{grone} for partial matrices whose corresponding graph is chordal.
-The procedure provided in \cite{grone} was improved in \cite{smith} by giving a way to calculate whole blocks of the completion at once.
+The procedure provided in \cite{grone} was improved upon in \cite{smith} by giving a way to calculate whole blocks of the completion at once.
 For positive definite matrices the resulting completion is singled out by two uniqueness results.

### 2021-12-03 paper/paper_math.tex
Submitted version.

@@
 
-\begin{abstract} Positive semidefinite Hermitian matrices that are not fully specified can be completed to a positive semidefinite Hermitian matrix provided their underlying graph is chordal.
+\begin{abstract} Positive semidefinite Hermitian matrices that are not fully specified can be completed provided their underlying graph is chordal.
 If the matrix is positive definite the completion can be uniquely characterized as the matrix that maximizes the determinant, or as the matrix whose inverse has zeroes in those places that were undetermined in the original matrix.

### 2021-12-03 paper/paper_math.tex
Encoding changes

@@
 	\bibitem{horn} R. A. Horn, Basic  Properties  of  the  Schur  Complement. In Fuzhen Zhang (Ed.), The Schur complement and its application (17--46), Springer, 2005.
-	\bibitem{schur} J. Schur, Über Potenzreihen, die im Innern des Einheitskreises beschränkt sind, Journal für die reine und angewandte Mathematik 147 (1917) 205 -- 232.
-	\bibitem{bana} T. Banachiewicz, Zur Berechnung der Determinanten, wie auch der Inversen, und zur darauf basierten Auflösung der Systeme linearer Gleichungen, Acta Astronom. S\'er. C 3 (1937) 41--67.	
+	\bibitem{schur} J. Schur, \"Uber Potenzreihen, die im Innern des Einheitskreises beschr\"ankt sind, Journal für die reine und angewandte Mathematik 147 (1917) 205 -- 232.
+	\bibitem{bana} T.
+Banachiewicz, Zur Berechnung der Determinanten, wie auch der Inversen, und zur darauf basierten Aufl\"osung der Systeme linearer Gleichungen, Acta Astronom.
+S\'er.
+C 3 (1937) 41--67.
 	\bibitem{frazer} R. Frazer, W. Duncan, A. Collar, Elementary Matrices And Some Applications To Dynamics And Differential Equations, Cambridge University Press, 1938.
@@
 https://dx.doi.org/10.1137/0117041
-	\bibitem{otherpaper} O. Dreyer, H. Köhler, T. Streuer, Completing correlation matrices, arXiv:2111.12640 
+	\bibitem{otherpaper} O. Dreyer, H. K\"ohler, T. Streuer, Completing correlation matrices, arXiv:2111.12640 
 \end{thebibliography}
